### 2022-12-21 thesis_template/paper0904.tex
added template for diabetes lit review, not complete yet

\documentclass{article}
%\documentclass[twocolumn]{article}
%\documentclass[onecolumn]{article}
% \usepackage{scrtime} % for \thistime (this package MUST be listed first!)
\DeclareUnicodeCharacter{0301}{\'{e}}
\usepackage{times}
\usepackage{graphicx}
\usepackage{float}
\usepackage[margin=0.75in]{geometry}
\usepackage{fancyhdr}
\usepackage{caption}
\usepackage{notoccite}
\usepackage{pgfplotstable}
%\usepackage[round]{natbib}
%\setcitestyle{aysep={}} %removes the comma between the author and year in citations
%\usepackage{underscore}
\usepackage{pdfpages}
\usepackage{xcolor,colortbl}%for changing cell colour
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{capt-of}
\pagestyle{fancy}
\setlength{\headheight}{15.2pt}
\setlength{\headsep}{13 pt}
\setlength{\parindent}{28 pt}
\setlength{\parskip}{12 pt}
\pagestyle{fancyplain}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
% \usepackage{color,amsmath,amssymb,amsthm,mathrsfs,amsfonts,dsfont}
\usepackage{xspace}
\usepackage{tikz-cd}
\usepackage{tikz}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{calc, arrows}
\usetikzlibrary{external}
\usepackage{pgfplots}
\pgfplotsset{layers/my layer set/.define layer set={background,main,foreground}{},
        set layers=my layer set,}

\usepackage{listings}
\usepackage{xcolor}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

%for code
\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\ttfamily\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}

\lstset{style=mystyle}
% \usetikzlibrary{pgfplots.clickable}
% \usepgfplotslibrary{clickable}
% tables
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
% figs
%\usepackage{subfig}% http://ctan.org/pkg/subfig
\usepackage{subcaption}
%\newsubfloat{figure}% Allow sub-figures
\usepackage{caption}%lable fig caption as fig
%\captionsetup[subfigure]{labelfont=bf, justification=raggedright, labelformat=empty} %no caption label
\usepackage{stackengine} %places caption inside figure?
\captionsetup{subrefformat=parens} %when you reference the subcaption it will be (a) for example %{labelfont={color=blue}}
%\captionsetup[subfigure]{labelsep=colon}

% \usepackage{acronym}
% \usepackage{lineno}%for line numbers
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BIBLIOGRAPHY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[backend=biber, giveninits=true, doi=false, isbn=false, natbib=true, url=true, eprint=false, style=authoryear-comp, sorting=nyt, sortcites=ynt, maxcitenames=2, maxbibnames=10, minbibnames = 10, uniquename=false, uniquelist=false, dashed=false]{biblatex} % can change the maxbibnames to cut long author lists to specified length followed by et al., currently set to 99.

%% bibliography for each chapter...
\DeclareFieldFormat[article,inbook,incollection,inproceedings,patent,thesis,unpublished]{title}{#1\isdot} % removes quotes around title
\renewbibmacro*{volume+number+eid}{%
	\printfield{volume}%
	%  \setunit*{\adddot}% DELETED
	\printfield{number}%
	\setunit{\space}%
	\printfield{eid}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
%\renewcommand*{\newunitpunct}{\space} % remove period after date, but I like it. 
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}} % this remove the "In: Journal Name" from articles in the bibliography, which happens with the ynt 
\renewbibmacro*{note+pages}{%
	\printfield{note}%
	\setunit{,\space}% could add punctuation here for after volume
	\printfield{pages}%
	\newunit}    
\DefineBibliographyStrings{english}{% clears the pp from pages
	page = {\ifbibliography{}{\adddot}},
	pages = {\ifbibliography{}{\adddot}},
} 
\DeclareFieldFormat{journaltitle}{#1\isdot}
\renewcommand*{\revsdnamepunct}{}%remove comma between last name and first name
\DeclareNameAlias{sortname}{family-given}
% \DeclareNameAlias{sortname}{last-first}
\renewcommand*{\nameyeardelim}{\addspace} % remove comma in text between name and date
\addbibresource{ABC1.bib} % The filename of the bibliography
\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\renewrobustcmd*{\bibinitperiod}{}
% you'll have to play with the citation styles to resemble the standard in your field, or just leave them as is here. 
% or, if there is a bst file you like, just get rid of all this biblatex stuff and go back to bibtex. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% generally hyperref needs to be loaded last
\usepackage[hidelinks,colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=darkbrown]{hyperref}
%\usepackage[hidelinks,colorlinks=false,citecolor=blue,urlcolor=darkbrown]{hyperref}
\tikzexternalize

\lhead{Type 1 Diabetes Review} %This needs to change
\rhead{Bianca Flaim, Oluwalonimi Ayekun, and Alexander Turco}
\title{\sc Examining the Molecular Basis of Type 1 Diabetes}
\author{\sc Bianca Flaim, Oluwalonimi Ayekun, and Alexander Turco}

\providecommand{\figref}[1]{(Figure \ref{#1})}  %what?
\providecommand{\tabref}[1]{(Table \ref{#1})}  %what?
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\newcommand{\seg}{\texttt{Seg}\xspace}
\newcommand{\ecoli}{\mbox{\textit{E.\,coli}}\xspace}
\newcommand{\sclong}{\textit{Saccharomyces cerevisiae}\xspace}
\newcommand{\scshrt}{\mbox{\textit{S.\,cerevisiae}}\xspace}
\newcommand{\sce}{\mbox{\textit{S.\,cerevisiae}}\xspace}
\newcommand{\hslong}{\textit{Homo sapiens}\xspace}
\newcommand{\hsshrt}{\mbox{\textit{H.\,sapiens}}\xspace}
\newcommand{\hse}{\mbox{\textit{H.\,sapiens}}\xspace}
\newcommand{\celong}{\textit{Caenorhabditis elegans}\xspace}
\newcommand{\ceshrt}{\mbox{\textit{C.\,elegans}}\xspace}
\newcommand{\dmlong}{\textit{Drosophila melanogaster}\xspace}
\newcommand{\dmshrt}{\mbox{\textit{D.\,melanogaster}}\xspace}
\newcommand{\atlong}{\textit{Arabidopsis thaliana}\xspace}
\newcommand{\atshrt}{\mbox{\textit{A.\,thaliana}}\xspace}
\newcommand{\pflong}{\textit{Plasmodium falciparum}\xspace}
\newcommand{\pfshrt}{\mbox{\textit{P.\,falciparum}}\xspace}

%%%%BIBLIOGRAPHY

%Supplementary File Table Numbers:
\newcommand{\expdata}{S1\xspace}
\newcommand{\seqdata}{S2\xspace}
\newcommand{\protdata}{S3\xspace}
\newcommand{\blast}{S4\xspace}
\newcommand{\tabvar}{S5\xspace}
%Supplementary File Fig Numbers:

\newcommand{\expcor}{S1\xspace}
\newcommand{\expdistATCC}{S3\xspace}
\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\newcommand{\beginsupplement}{%
	\setcounter{table}{0}
	\renewcommand{\thetable}{S\arabic{table}}%    %thetable references table counter 
	\setcounter{figure}{0}
	\renewcommand{\thefigure}{S\arabic{figure}}%
	\setcounter{equation}{0}
	\renewcommand{\theequation}{S\arabic{equation}}%

}
\renewcommand{\thesection}{}
\renewcommand{\thesubsection}{}
\renewcommand{\thesubsubsection}{}
\usepackage{setspace}
%adjust spacing
\doublespacing
\usepackage{titlesec}

\titlespacing\section{0pt}{12pt plus 2pt minus 2pt}{0pt plus 1pt minus 1pt}
\titlespacing\subsection{0pt}{12pt plus 2pt minus 2pt}{0pt plus 1pt minus 1pt}
\titlespacing\subsubsection{0pt}{12pt plus 2pt minus 2pt}{0pt plus 1pt minus 1pt}

% below 3 lines will put ALL table captions at the top...not sure if
% this is what we want but it is good enough for now

% \usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

        \definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
        \definecolor{darkbrown}{rgb}{1.0, 0.56, 0.24}
        \colorlet{darkcol}{black!30!white}
        \colorlet{lightcol}{black!10!white}
        \definecolor{txtcol}{HTML}{F40000}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\widowpenalty10000
\clubpenalty10000

%\linenumbers %for line numbers
\onecolumn
%\twocolumn[  
%       \begin{@twocolumnfalse}
%               \begin{center}
                        \maketitle
%               \end{center}
%                       \bigskip

\thispagestyle{empty}
\noindent \textsuperscript{1} Department of Biology, McMaster University, Hamilton, ON, Canada

\newpage
\tableofcontents
\newpage

\section{Literature Review/Proposal}

\subsection{What is Type 1 Diabetes?}
Over the last 70 years, developed countries around the world have seen a steady increase in the incidence of type 1
diabetes (T1D) (Patterson et al. 2012; Todd 2010). It is generally thought that T1D is due to the immune-mediated destruction of
pancreatic $\beta$-cells which are responsible for the production of insulin (Atkinson, G S Eisenbarth, and Michels 2014; Todd 2010).
Autoantibodies specifically targeting $\beta$-cells, as well as macrophage and lymphocyte infiltration of the pancreas are responsible
for the destruction of insulin-producing $\beta$-cells (Van den Driessche et al. 2009). These $\beta$-cells are found within the islets of Langerhans and they have the ability to sense glucose and subsequently release insulin to maintain healthy blood-glucose levels (Bluestone, Herold, and G Eisenbarth 2010). Although it is known to be a common chronic disease in children, diagnosis of
T1D can occur at any age (Atkinson, G S Eisenbarth, and Michels 2014). Symptoms associated with disease onset, especially in
children and adolescents, are typically polydispia (extreme thirst), polyphagia (extreme hunger), polyuria (excessive urination),
and hyperglycaemia (high blood sugar) (Atkinson, G S Eisenbarth, and Michels 2014). Research surrounding T1D tends to
revolve around the search for environmental triggers, as well as genetic factors which contribute to the disease (Todd 2010).
Rapid advancements in both molecular biology and genetic sequencing methods have contributed to an increase in papers which
examine the genetic basis of the disease. It is believed that the nature component (genetic factors, molecules, cells) of T1D will
provide insight into disease mechanisms, and at the same time help researchers to identify environmental factors which play a
role in the development of the disease (Todd 2010).

\subsection{Insulin Hormone Production and Release}
The insulin hormone is a peptide hormone encoded on INS gene, found on chromosome 11 (cluster p15.5) (Ohneda,
Ee, and German 2000; Tokarz, MacDonald, and Klip 2018). Gene expression is regulated by several enhancer elements found
upstream of the INS gene, including IDX1, MafA, NeuroD1 (Tokarz, MacDonald, and Klip 2018). As insulin is a key hormone
involved in the feedback loop to establish blood glucose homeostasis, glucose levels also influence insulin production (Ohneda,
Ee, and German 2000; Vranic, Hollenberg, and Steiner 1985). Transcription from the INS gene yields a preproinsulin mRNA
molecule, which exits the $\beta$-cell’s nucleus and is translated by ribosomes situated within the rough endoplasmic reticulum
(RER) to produce a preproinsulin polypeptide (Tokarz, MacDonald, and Klip 2018). After translation, a series of posttrans-
lational modifications are required, beginning with removal of the signal sequence within the RER (Tokarz, MacDonald, and
Klip 2018). Signal removal allows polypeptide (now proinsulin) trafficking to the trans-Golgi network, where the C-peptide
is proteolytically cleaved by the prohormone convertases PC1/3 and PC2 (Klimek et al. 2009; Tokarz, MacDonald, and Klip
2018). Proinsulin is then appropriately folded and stabilized through the formation of disulfide bonds, yielding a mature insulin
protein (Mathieu, Gillard, and Benhalima 2017; Tokarz, MacDonald, and Klip 2018). Insulin is then packaged into granules
and trafficked along the cell’s actin microtubule network where they are stored within the $\beta$-cell’s plasma membrane, prepared
for secretion once the cell receives the appropriate signal to release insulin (Tokarz, MacDonald, and Klip 2018).

Glucose entry in the $\beta$-cell triggers a cascade of intracellular events which terminates in insulin release. Once blood glu-
cose rises typically above 100 mg/dL, glucose begins to enter the $\beta$-cell’s cytoplasm via membrane GLUT1 channels (Tokarz,
MacDonald, and Klip 2018; Vranic, Hollenberg, and Steiner 1985). Glucokinase then phosphorylates the glucose, trapping it
within the cell and beginning the process of glycolysis, resulting in the production of adenosine triphosphate (ATP) and pyru-
vate (Tokarz, MacDonald, and Klip 2018). Pyruvate then undergoes subsequent metabolic steps, including entering the Krebs
cycle and the electron transport chain within the mitochondria, resulting in the production of lots of ATP. Within the cytoplasm,
the increased ATP/adenosine diphosphate (ADP) ratio is detected, resulting in the closure of ATP-sensitive K+ ion channels
(Tokarz, MacDonald, and Klip 2018). Since the potassium ions ca no longer flow out of the cell, the closure of these channels
results in a more positive membrane potential, leading to depolarization of the cell once the membrane potential reaches –50
millivolts (mV) (Tokarz, MacDonald, and Klip 2018). Depolarization of the cell activates voltage-dependent Na+ and Ca2+
channels, allowing the influx of sodium and calcium ions, further increasing membrane potential (Tokarz, MacDonald, and Klip
2018). The influx of calcium ions triggers SNARE proteins, Munc18, and syntaxin proteins to initiate insulin vesicle fusion and
release into the bloodstream (Tokarz, MacDonald, and Klip 2018). As insulin is released and plasma glucose levels begin to
drop, the signal is terminated as the SERCA pump removes calcium ions from the $\beta$-cell’s cytoplasm (Tokarz, MacDonald, and
Klip 2018). Additionally, the signal is terminated as granule exocytosis requires ATP hydrolysis, thus reducing the cytosolic
ATP:ADP ratio, and triggering the reopening of ATP-sensitive K+ channels (Tokarz, MacDonald, and Klip 2018). Allowing
potassium ions to flow out of the cell reduces the membrane potential, resulting in the closure of voltage-gated Na+ and Ca2+
channels. The process of insulin release has been described as a biphasic process, with an initial burst as the cell initially depo-
larizes and releases all primed insulin granules within the cell membrane, and a second, prolonged phase as the cell’s network
reorganizes itself to recruit more insulin granules to the plasma membrane for subsequent exocytosis (Tokarz, MacDonald,
and Klip 2018). Within an islet of Langerhans, which contains many $\beta$-cells, which are connected by gap junctions, allowing
rapid paracellular cell-to-cell communication, and resulting in a large, rapid, release of insulin from the pancreas when needed
(In’t Veld and Marichal 2010; Tokarz, MacDonald, and Klip 2018).

\subsection{Insulin-Insulin Receptor Binding Mediated Glucose Storage and Uptake}
Insulin binds to the insulin receptor (IR) expressed on the cell membrane of its target cells. The IR is a disulphide-linked
receptor tyrosine kinase homodimer comprised of two $\alpha$-subunits and two $\beta$-subunits. After insulin binding, the IR undergoes a conformational change bringing the two IR monomers closer and allowing for autophosphorylation (De Meyts 2016). During
autophosphorylation, each IR monomer adds a phosphate group to the $\beta$-unit of the other IR monomer resulting in the activation
of the IR receptor. Once activated, the IR becomes capable of activating and binding to the adaptor protein insulin-receptor
substrate 1 (IRS-1). IRS-1 acts as a docking site for proteins with Src-homology-2 domains, such as the p85 subunit of PI
(3)K which helps facilitate the recruitment PI (3)K to the plasma membrane (Lizcano and Alessi 2002; Sun et al. 1991). Upon
being recruited to the plasma membrane, PI (3)K catalyzes the conversion of phosphatidylinositol (4,5)-biphosphate (PI (4,5)
P2) into phosphatidylinositol (3,4,5) triphosphate (PI (3,4,5) P3). PI (3,4,5) P3 can phosphorylate the AKT (also known as
protein kinase B) (Sun et al. 1991). The increased concentration of PI (3,4,5) P3 molecules along the cell membrane facilitates
the recruitment of AKT to the plasma membrane. Once at the plasma membrane, AKT is phosphorylated by two other protein
kinases, PDK1 at residue Thre308 and MTORC at residue Ser473. The phosphorylation of the protein at these two residues
allows for its activation (Leto and Saltiel 2012; Menting et al. 2013). Upon its activation, AKT dissociates from the membrane
and begins to interact with number of different effector proteins within the cytosol and nucleus, including GSK3 (Lizcano and
Alessi 2002). GSK3 is a serine/threonine protein kinase that can inhibit the activity of glycogen synthase, a key enzyme in the
process of glycogenesis. When AKT phosphorylates GSK3 it inhibits the activity of this kinase, therefore increasing glycogen
synthase and the rate of glucose-to-glycogen conversion. This is how insulin increases glucose storage inside of muscle and fat
cells.

In addition to glucose storage, insulin also helps facilitate the process of GLUT-4 translocation to the cell membrane
in muscle and adipose (fat) cells. GLUT-4 is one of the several GLUT proteins responsible for transporting glucose across
the blood vessel into cells in the human body (Li et al. 2019). When insulin is not bound to the IR, about 80% of GLUT-4
remains sequestered in specialized, intracellular organelles known as GLUT-4 storage vesicles (GSVs) while the other 20%
cycle slowly between the GSVs and plasma membrane (Fazakerley, Koumanov, and Holman 2022). Within 5-10 minutes of
insulin-mediated intracellular signalling, the number of GLUT-4 receptors expressed along the surface of adipose and fat cells
increases by over 50% (Fazakerley, Koumanov, and Holman 2022; Stöckli, Fazakerley, and James 2011). Insulin increases
the amount of GLUT-4 at the plasma membrane primarily by promoting the fusion of GSVs with the plasma membrane (also
known as GSVs exocytosis) (Leto and Saltiel 2012). Like with glucose storage, AKT acts as a central mediator of this process
through its intracellular interactions with several important molecules and proteins involved in GSVs exocytosis. For instance,
AKT is capable of phosphorylating both RAB-GAP AS160 (also referred to asTBC1D4) and the RAL-GAP complex, each of
which have important roles in GLUT-4 vesicle trafficking and targeting to the plasma membrane (De Meyts 2016; Leto and
Saltiel 2012). Along with RAB-GAP AS160 and the RAL-GAP complex, AKT has also been shown to interact with SNARE
regulatory proteins, SYNIP and CDP138, which are necessary for the fusion of GSVs to the plasma membrane.

In addition to the PI (3)K-AKT pathway, however, insulin is also capable of promoting GLUT-4 translocation via another
intracellular signalling pathway. This pathway is mediated by the adaptor protein APS, which binds with high affinity to
the insulin-receptor (Yu et al. 2007). After APS is phosphorylated by the IR, it recruits a complex consisting of the proto-
oncogene c-CBL and the-CBL associated protein (CAP) facilitating the IR-catalyzed Tyr phosphorylation of c-CBL. Following
phosphorylation by the IR, c-CBL becomes capable of interacting with the adaptor protein CRK which is bound to the GEF
(Guanine-Exchange-Factor) C3G. This interaction allows C3G to initiate the activation of TC10, an RHO-family GTP-Ase,
localized in the lipid rafts of the plasma membrane (Leto and Saltiel 2012). TC10 is then able to interact with its’ effector
protein called PIST 1, which then binds to and directs the cleavage of the protein TUG. In basal conditions (under low insulin
concentration), TUG binds to GLUT-4 in GSVs preventing the trafficking of these GSVs, and thus GLUT-4 translocation, to the
plasma membrane. Insulin-induced cleavage of the TUG protein stimulates the dissociation of protein from GLUT-4 enabling
the trafficking of GSVs to the plasma membrane (Li et al. 2019; Yu et al. 2007). In the absence of this step, GSVs regulated
trafficking and GLUT-4 translocation is inhibited in adipocyte cells (Chang, Chiang, and Saltiel 2004). In addition to facilitating
the cleavage of TUG, TC10 also interacts with the protein EXO70 which belongs to the exocyst complex and is involved in the
docking of secretory vesicles to the plasma membrane.

\subsection{Type 1 Diabetes Symptom Pathophysiology and Possible Complications}
T1D is said to proceed in 3 phases – two presymptomatic phases, followed by a symptomatic phase that increases in
severity as the disease progresses (Katsarou et al. 2017; Norris, Johnson, and Stene 2020). The first phase is characterized
by initial immune infiltration of the pancreas by macrophages, dendritic cells, CD8+ and CD4+ T-cells, and autoantibodies
specific against $\beta$-cells, resulting in the rapid loss of functional $\beta$-cells (Katsarou et al. 2017). This stage is followed by the initial development of hyperglycemia, which may present as asymptomatic at first as the pancreas retains some functionality (Katsarou et al. 2017). As immune destruction continues, $\beta$-cell functionality dramatically decreases, resulting in more severe hyperglycemia and the development of symptoms (Katsarou et al. 2017). Primary symptoms include polyuria, polydipsia, and polyphagia (R A Guthrie and D W Guthrie 2004). As glucose accumulates within the bloodstream, the kidneys attempt to
rectify the imbalance by excreting more glucose in the urine (R A Guthrie and D W Guthrie 2004). As the nephron excretes
glucose into the urine, water follows from the bloodstream into the urine, resulting in a greater volume of urine and the primary
symptom of polyuria (R A Guthrie and D W Guthrie 2004). Increased urination in attempts to reduce blood glucose results
in dehydration, so this causes the polydipsia as the body’s mechanism to increase water intake and restore hydration and
water homeostasis (R A Guthrie and D W Guthrie 2004). The final primary symptom, polyphagia, is increased hunger due to
reduced fuel available to cells (R A Guthrie and D W Guthrie 2004). Primary complications include nephropathy, neuropathy,
retinopathy, risk of cardiovascular disease and peripheral artery disease, and risk of cerebrovascular disease (Katsarou et al.
2017; McGill and Levitsky 2016; Melendez-Ramirez, Richards, and Cefalu 2010; Wherrett et al. 2018). These conditions
are largely related to the severity and length of hyperglycemic exposure, especially due to the increased stress on the kidneys
due to increased glucose excretion. As hyperglycemia exposure increases, this can increase the risk of albuminuria, which
may progress to complete nephropathy (Melendez-Ramirez, Richards, and Cefalu 2010). Neuropathy may arise in the case of
hypoglycemia (in instances where too much insulin is injected), as the brain primarily relies on glucose as its fuel source, and
reduced glucose delivery to the brain may result in neuronal death and possible cognitive impairments and deficits (Katsarou
et al. 2017). Hyperglycemia is also linked to the production of microaneurysms in blood vessels feeding the retina, leading
to possible retinopathy and fluid leakage resulting in macular edema (Melendez-Ramirez, Richards, and Cefalu 2010). Lastly,
hyperglycemia can trigger a severe complication known as diabetic ketoacidosis (DKA). Hyperglycemia causes the production
of ketones as an alternate source of fuel when insulin is critically low, and cells cannot access glucose (Cashen and Petersen
2019). In this case, ketogenesis begins, and can result in the acidification of the blood, and can also result in cerebral edema
and hypokalemia which can become fatal in severe cases (Cashen and Petersen 2019).

\subsection{Mutations associated with Type 1 Diabetes}
T1D is a complex disease, meaning it can be influenced by a variety of environmental and genetic risk factors (Brorsson
et al. 2010). The concordance of T1D among monozygotic and dizygotic twins, as well as numerous studies that have identified
genes associated with T1D, provide good evidence for a strong genetic determinant of disease (Erlich et al. 2008). For over
thirty years, it has been known that the human leukocyte antigen (HLA) region on chromosome 6 contributes to the risk of
developing T1D (Brorsson et al. 2010). The HLA region contains genes which code for proteins that have immune function,
particularly in the regulation of the immune response (Choo 2007). More specifically, it has been found that certain alleles at
the HLA-DRB1, HLA-DQA1, and HLA-DQB1 loci are more strongly associated with T1D than others (Erlich et al. 2008).
Although studies have identified these specific alleles, the consensus remains that more work must be done to further identify
causative variants and their DNA profile (Brorsson et al. 2010). The complexity of T1D makes it extremely difficult to narrow
down the cause to one single gene, but there are a very small number of diabetes cases which do in fact result from single
gene mutations (Yang and Chan 2016). Two major forms of monogenic diabetes are Neonatal diabetes mellitus (NDM) and
maturity-onset diabetes of the young (MODY) (Yang and Chan 2016). These monogenic forms of diabetes are typically the
result of mutations in genes which code for proteins that reduce $\beta$-cell function (Steck and Winter 2011). It is interesting to
note that the single gene mutations found to cause NDM and MODY, have also been found to be associated with T1D (Yang
and Chan 2016). Two examples of monogenic diabetes genes which have also been found to be associated with T1D are the
preproinsulin gene (INS), and the Gli-similar 3 gene (GLIS3) (Yang and Chan 2016). The GLIS3 gene codes for a transcription
factor which regulates islet development and insulin gene transcription, and the INS gene encodes the peptide hormone insulin
(Steck and Winter 2011; Yang and Chan 2016). Single nucleotide polymorphisms (SNPs) are typically what affect the normal
function of both genes, but it has also been found that allelic variants of the genes lead to both monogenic diabetes and T1D
(Yang and Chan 2016).

\subsection{Type 1 Diabetes Treatments and Health Considerations}
As the $\beta$-cell functionality is destroyed due to immune destruction, functionality cannot be restored. Thus, the standard
of care involves continuous glucose monitoring and regulation of blood sugar levels via insulin injections (McGill and Levitsky
2016). Many patients opt for insulin pumps, otherwise known as a closed-loop pancreas system (Wherrett et al. 2018). This
system is an artificial system that features an insulin pump, and a glucose sensor which can detect when insulin should be
secreted. Some of these systems may infuse just insulin, others also feature a glucagon pump to increase blood sugar in case
too much insulin was secreted (Wherrett et al. 2018). Alternatively, patients may opt for standard insulin injections. A key
component of T1D treatment is patient education on their condition and the body’s natural fluctuations of blood sugar levels
throughout the day. For example, it is important for patients to understand that prolonged fasting, illness, or intense exercise can
reduce blood sugar levels and naturally cause a hypoglycemic crisis (McGill and Levitsky 2016). In the event of hypoglycemia,
it is important to detect as soon as possible and consume a small amount of sugar (15 grams typically recommended, available
in juice, candies, or glucose gels) (Katsarou et al. 2017). Lifestyle changes such as eating snacks prior to bedtime and exercising
earlier in the day can be beneficial in preventing nocturnal hypoglycemia (McGill and Levitsky 2016). Due to the number of
potential complications which may arise from T1D, physicians should also monitor patients for other comorbidities including
other autoimmune conditions, kidney failure, dyslipidemia, retinopathy, and other hormonal disease (Katsarou et al. 2017).
Further, patients are encouraged to consult dieticians to develop healthy meal plans to regulate blood sugar levels and join
community support groups (Katsarou et al. 2017). T1D patients are at increased risk of mental health issues such as anxiety,
depression, and eating disorders due to the perceived lack of autonomy and increased stress load due to their disease (Katsarou
et al. 2017).

\clearpage\newpage
\section{References}

%%%FIGURES%%%%%

%%%%PRINTING BIBLIOGRAPHY%%%%
\nocite{*}
\printbibliography[heading=none, sorting=nyt]
\newpage

\end{document}

### 2023-01-08 thesis_template/paper0904.tex
adding citations in bibtex

@@
 
-\lhead{Type 1 Diabetes Review} %This needs to change
-\rhead{Bianca Flaim, Oluwalonimi Ayekun, and Alexander Turco}
-\title{\sc Examining the Molecular Basis of Type 1 Diabetes}
-\author{\sc Bianca Flaim, Oluwalonimi Ayekun, and Alexander Turco}
+\lhead{Cells at War: Game-Based Learning} %This needs to change
+\rhead{Bianca Flaim, Alexander Turco}
+\title{\sc Cells at War: The Playfulness of Game-Based Learning}
+\author{\sc Bianca Flaim, Alexander Turco and Rosa da Silva}
 
@@
 
+\section{Abstract}
+
+\newpage
+
 \section{Literature Review/Proposal}
 
-\subsection{What is Type 1 Diabetes?}
-Over the last 70 years, developed countries around the world have seen a steady increase in the incidence of type 1
-diabetes (T1D) (Patterson et al. 2012; Todd 2010). It is generally thought that T1D is due to the immune-mediated destruction of
-pancreatic $\beta$-cells which are responsible for the production of insulin (Atkinson, G S Eisenbarth, and Michels 2014; Todd 2010).
-Autoantibodies specifically targeting $\beta$-cells, as well as macrophage and lymphocyte infiltration of the pancreas are responsible
-for the destruction of insulin-producing $\beta$-cells (Van den Driessche et al.
-2009).
-These $\beta$-cells are found within the islets of Langerhans and they have the ability to sense glucose and subsequently release insulin to maintain healthy blood-glucose levels (Bluestone, Herold, and G Eisenbarth 2010).
-Although it is known to be a common chronic disease in children, diagnosis of
-T1D can occur at any age (Atkinson, G S Eisenbarth, and Michels 2014). Symptoms associated with disease onset, especially in
-children and adolescents, are typically polydispia (extreme thirst), polyphagia (extreme hunger), polyuria (excessive urination),
-and hyperglycaemia (high blood sugar) (Atkinson, G S Eisenbarth, and Michels 2014). Research surrounding T1D tends to
-revolve around the search for environmental triggers, as well as genetic factors which contribute to the disease (Todd 2010).
-Rapid advancements in both molecular biology and genetic sequencing methods have contributed to an increase in papers which
-examine the genetic basis of the disease. It is believed that the nature component (genetic factors, molecules, cells) of T1D will
-provide insight into disease mechanisms, and at the same time help researchers to identify environmental factors which play a
-role in the development of the disease (Todd 2010).
+\subsection{Introduction - Maybe move to Abstract, it encompasses everything}
+After many years of trying to adapt to educational innovations of the 21st Century, educators will soon be looking into the not too distant future and considering what modern learning in post-secondary education will look like in the 22nd Century.
+Technological advances will continue to transform the landscape of the classroom and the use of digital games, although a source of controversy and debate, will prove to be highly effective instructional tools that can promote problem solving skills and critical thinking.
+Their use has already started to change the structure of the class in the elementary classroom as they serve to energize and engage students in their learning and improve learning, creativity and imagination.
+The traditional construct of the classroom has started to evolve, embracing a new way in which knowledge, skills and attitudes can be acquired.
+Cells at War is a collection of scientific video games developed through interdisciplinary efforts that reinforces the notion that video games are highly effective tools used to enhance instructional methods.
+(Not only enhancing but teaching and engaging, ask rosa).
+– I don’t know if I should mention the game but I like this line.
 
-\subsection{Insulin Hormone Production and Release}
-The insulin hormone is a peptide hormone encoded on INS gene, found on chromosome 11 (cluster p15.5) (Ohneda,
-Ee, and German 2000; Tokarz, MacDonald, and Klip 2018). Gene expression is regulated by several enhancer elements found
-upstream of the INS gene, including IDX1, MafA, NeuroD1 (Tokarz, MacDonald, and Klip 2018). As insulin is a key hormone
-involved in the feedback loop to establish blood glucose homeostasis, glucose levels also influence insulin production (Ohneda,
-Ee, and German 2000; Vranic, Hollenberg, and Steiner 1985). Transcription from the INS gene yields a preproinsulin mRNA
-molecule, which exits the $\beta$-cell’s nucleus and is translated by ribosomes situated within the rough endoplasmic reticulum
-(RER) to produce a preproinsulin polypeptide (Tokarz, MacDonald, and Klip 2018). After translation, a series of posttrans-
-lational modifications are required, beginning with removal of the signal sequence within the RER (Tokarz, MacDonald, and
-Klip 2018). Signal removal allows polypeptide (now proinsulin) trafficking to the trans-Golgi network, where the C-peptide
-is proteolytically cleaved by the prohormone convertases PC1/3 and PC2 (Klimek et al. 2009; Tokarz, MacDonald, and Klip
-2018). Proinsulin is then appropriately folded and stabilized through the formation of disulfide bonds, yielding a mature insulin
-protein (Mathieu, Gillard, and Benhalima 2017; Tokarz, MacDonald, and Klip 2018). Insulin is then packaged into granules
-and trafficked along the cell’s actin microtubule network where they are stored within the $\beta$-cell’s plasma membrane, prepared
-for secretion once the cell receives the appropriate signal to release insulin (Tokarz, MacDonald, and Klip 2018).
+\subsection{Children and Play}
 
-Glucose entry in the $\beta$-cell triggers a cascade of intracellular events which terminates in insulin release. Once blood glu-
-cose rises typically above 100 mg/dL, glucose begins to enter the $\beta$-cell’s cytoplasm via membrane GLUT1 channels (Tokarz,
-MacDonald, and Klip 2018; Vranic, Hollenberg, and Steiner 1985). Glucokinase then phosphorylates the glucose, trapping it
-within the cell and beginning the process of glycolysis, resulting in the production of adenosine triphosphate (ATP) and pyru-
-vate (Tokarz, MacDonald, and Klip 2018). Pyruvate then undergoes subsequent metabolic steps, including entering the Krebs
-cycle and the electron transport chain within the mitochondria, resulting in the production of lots of ATP. Within the cytoplasm,
-the increased ATP/adenosine diphosphate (ADP) ratio is detected, resulting in the closure of ATP-sensitive K+ ion channels
-(Tokarz, MacDonald, and Klip 2018). Since the potassium ions ca no longer flow out of the cell, the closure of these channels
-results in a more positive membrane potential, leading to depolarization of the cell once the membrane potential reaches –50
-millivolts (mV) (Tokarz, MacDonald, and Klip 2018). Depolarization of the cell activates voltage-dependent Na+ and Ca2+
-channels, allowing the influx of sodium and calcium ions, further increasing membrane potential (Tokarz, MacDonald, and Klip
-2018). The influx of calcium ions triggers SNARE proteins, Munc18, and syntaxin proteins to initiate insulin vesicle fusion and
-release into the bloodstream (Tokarz, MacDonald, and Klip 2018). As insulin is released and plasma glucose levels begin to
-drop, the signal is terminated as the SERCA pump removes calcium ions from the $\beta$-cell’s cytoplasm (Tokarz, MacDonald, and
-Klip 2018). Additionally, the signal is terminated as granule exocytosis requires ATP hydrolysis, thus reducing the cytosolic
-ATP:ADP ratio, and triggering the reopening of ATP-sensitive K+ channels (Tokarz, MacDonald, and Klip 2018). Allowing
-potassium ions to flow out of the cell reduces the membrane potential, resulting in the closure of voltage-gated Na+ and Ca2+
-channels. The process of insulin release has been described as a biphasic process, with an initial burst as the cell initially depo-
-larizes and releases all primed insulin granules within the cell membrane, and a second, prolonged phase as the cell’s network
-reorganizes itself to recruit more insulin granules to the plasma membrane for subsequent exocytosis (Tokarz, MacDonald,
-and Klip 2018). Within an islet of Langerhans, which contains many $\beta$-cells, which are connected by gap junctions, allowing
-rapid paracellular cell-to-cell communication, and resulting in a large, rapid, release of insulin from the pancreas when needed
-(In’t Veld and Marichal 2010; Tokarz, MacDonald, and Klip 2018).
+Although technological ecosystems have been established in schools for some time, its landscape continues to be missing significant components to create a fully functioning ecosystem.
+One of these key components is that educators are reticent to embrace digital technology and gaming as the new norm in the classroom.
+Kindergarten classrooms in Ontario are bustling places of activity, reflective of active inquiry through play.
+Exploration and provocations drive their curiosity while simultaneously developing their oral language and critical thinking skills \citep{kindergartencurriculum}.
+Digital games on ipads are being used to develop literacy and numeracy skills.
+Some educators however, oppose the use of digitized games especially in higher education and have not fully embraced the notion that learning can actually take place through play.
+We believe that play is paramount for children in their formative years and educational theorists like Vygotsky affirm this.
+He was firm in his belief that play is the leading line of development during preschool years \citep{vygotsky1967play}.
+Unfortunately, in many classrooms, educators continue to feel more comfortable in engaging in antiquated and traditional methods of instruction that consist of a didactic approach to learning.
+The need to challenge current educational practices is paramount as educators are currently teaching students from the perspective of their own past rather than the realities of the students’ future.
+– I like this line, its powerful that’s why I highlighted it.
 
-\subsection{Insulin-Insulin Receptor Binding Mediated Glucose Storage and Uptake}
-Insulin binds to the insulin receptor (IR) expressed on the cell membrane of its target cells. The IR is a disulphide-linked
-receptor tyrosine kinase homodimer comprised of two $\alpha$-subunits and two $\beta$-subunits.
-After insulin binding, the IR undergoes a conformational change bringing the two IR monomers closer and allowing for autophosphorylation (De Meyts 2016).
-During
-autophosphorylation, each IR monomer adds a phosphate group to the $\beta$-unit of the other IR monomer resulting in the activation
-of the IR receptor. Once activated, the IR becomes capable of activating and binding to the adaptor protein insulin-receptor
-substrate 1 (IRS-1). IRS-1 acts as a docking site for proteins with Src-homology-2 domains, such as the p85 subunit of PI
-(3)K which helps facilitate the recruitment PI (3)K to the plasma membrane (Lizcano and Alessi 2002; Sun et al. 1991). Upon
-being recruited to the plasma membrane, PI (3)K catalyzes the conversion of phosphatidylinositol (4,5)-biphosphate (PI (4,5)
-P2) into phosphatidylinositol (3,4,5) triphosphate (PI (3,4,5) P3). PI (3,4,5) P3 can phosphorylate the AKT (also known as
-protein kinase B) (Sun et al. 1991). The increased concentration of PI (3,4,5) P3 molecules along the cell membrane facilitates
-the recruitment of AKT to the plasma membrane. Once at the plasma membrane, AKT is phosphorylated by two other protein
-kinases, PDK1 at residue Thre308 and MTORC at residue Ser473. The phosphorylation of the protein at these two residues
-allows for its activation (Leto and Saltiel 2012; Menting et al. 2013). Upon its activation, AKT dissociates from the membrane
-and begins to interact with number of different effector proteins within the cytosol and nucleus, including GSK3 (Lizcano and
-Alessi 2002). GSK3 is a serine/threonine protein kinase that can inhibit the activity of glycogen synthase, a key enzyme in the
-process of glycogenesis. When AKT phosphorylates GSK3 it inhibits the activity of this kinase, therefore increasing glycogen
-synthase and the rate of glucose-to-glycogen conversion. This is how insulin increases glucose storage inside of muscle and fat
-cells.
+\subsection{What Happens to Play Beyond the Formative Years?}
 
-In addition to glucose storage, insulin also helps facilitate the process of GLUT-4 translocation to the cell membrane
-in muscle and adipose (fat) cells. GLUT-4 is one of the several GLUT proteins responsible for transporting glucose across
-the blood vessel into cells in the human body (Li et al. 2019). When insulin is not bound to the IR, about 80% of GLUT-4
-remains sequestered in specialized, intracellular organelles known as GLUT-4 storage vesicles (GSVs) while the other 20%
-cycle slowly between the GSVs and plasma membrane (Fazakerley, Koumanov, and Holman 2022). Within 5-10 minutes of
-insulin-mediated intracellular signalling, the number of GLUT-4 receptors expressed along the surface of adipose and fat cells
-increases by over 50% (Fazakerley, Koumanov, and Holman 2022; Stöckli, Fazakerley, and James 2011). Insulin increases
-the amount of GLUT-4 at the plasma membrane primarily by promoting the fusion of GSVs with the plasma membrane (also
-known as GSVs exocytosis) (Leto and Saltiel 2012). Like with glucose storage, AKT acts as a central mediator of this process
-through its intracellular interactions with several important molecules and proteins involved in GSVs exocytosis. For instance,
-AKT is capable of phosphorylating both RAB-GAP AS160 (also referred to asTBC1D4) and the RAL-GAP complex, each of
-which have important roles in GLUT-4 vesicle trafficking and targeting to the plasma membrane (De Meyts 2016; Leto and
-Saltiel 2012). Along with RAB-GAP AS160 and the RAL-GAP complex, AKT has also been shown to interact with SNARE
-regulatory proteins, SYNIP and CDP138, which are necessary for the fusion of GSVs to the plasma membrane.
+The need to engage and motivate learners does not cease to be a need in a student’s educational career.
+However, what does become a priority in higher academic years are measurable student outcomes \citep{ball2012politics,shore2010beyond,leather2021pedagogy}.
+Tests and quizzes take precedence over the formative process of learning.
+Assessment of learning (tests and quizzes) dominates an educator’s instructional practices which in turn diminishes the value of the experiential learning process (assessment as learning).
+Assessment methods seem to be based on antiquated teaching practices, which have been influenced by our Victorian educational heritage that is rooted in a persistent view of fear of play \citep{wood2013play}.
+This rigidity in instructional norms appears to have taken hold in our classrooms which diminishes the value and importance of games in the classroom.
+\citet{reich2020failure} argues that schools are complex and conservative institutions that have little incentive to adopt or test novel approaches.
 
-In addition to the PI (3)K-AKT pathway, however, insulin is also capable of promoting GLUT-4 translocation via another
-intracellular signalling pathway. This pathway is mediated by the adaptor protein APS, which binds with high affinity to
-the insulin-receptor (Yu et al. 2007). After APS is phosphorylated by the IR, it recruits a complex consisting of the proto-
-oncogene c-CBL and the-CBL associated protein (CAP) facilitating the IR-catalyzed Tyr phosphorylation of c-CBL. Following
-phosphorylation by the IR, c-CBL becomes capable of interacting with the adaptor protein CRK which is bound to the GEF
-(Guanine-Exchange-Factor) C3G. This interaction allows C3G to initiate the activation of TC10, an RHO-family GTP-Ase,
-localized in the lipid rafts of the plasma membrane (Leto and Saltiel 2012). TC10 is then able to interact with its’ effector
-protein called PIST 1, which then binds to and directs the cleavage of the protein TUG. In basal conditions (under low insulin
-concentration), TUG binds to GLUT-4 in GSVs preventing the trafficking of these GSVs, and thus GLUT-4 translocation, to the
-plasma membrane. Insulin-induced cleavage of the TUG protein stimulates the dissociation of protein from GLUT-4 enabling
-the trafficking of GSVs to the plasma membrane (Li et al. 2019; Yu et al. 2007). In the absence of this step, GSVs regulated
-trafficking and GLUT-4 translocation is inhibited in adipocyte cells (Chang, Chiang, and Saltiel 2004). In addition to facilitating
-the cleavage of TUG, TC10 also interacts with the protein EXO70 which belongs to the exocyst complex and is involved in the
-docking of secretory vesicles to the plasma membrane.
+It can be argued that this fear of play is also due in large part to limited resources and training being provided to educators.
+Teachers require technological and pedagogical supports to develop their understanding of game based learning in the classroom.
+The expectation therefore to incorporate new and innovative teaching methodologies in the classroom is unrealistic without providing teachers with the appropriate training needed to do so.
+Incorporating game based learning in teacher practice is impacted by a lack of professional development opportunities that can systematically guide them in using games for teaching, learning, and assessment \citep{fishman2014empowering,ruggiero2013video,foster2020principles}.
+Might want to add a line here, don’t wanna end on quote
 
-\subsection{Type 1 Diabetes Symptom Pathophysiology and Possible Complications}
-T1D is said to proceed in 3 phases – two presymptomatic phases, followed by a symptomatic phase that increases in
-severity as the disease progresses (Katsarou et al. 2017; Norris, Johnson, and Stene 2020). The first phase is characterized
-by initial immune infiltration of the pancreas by macrophages, dendritic cells, CD8+ and CD4+ T-cells, and autoantibodies
-specific against $\beta$-cells, resulting in the rapid loss of functional $\beta$-cells (Katsarou et al.
-2017).
-This stage is followed by the initial development of hyperglycemia, which may present as asymptomatic at first as the pancreas retains some functionality (Katsarou et al.
-2017).
-As immune destruction continues, $\beta$-cell functionality dramatically decreases, resulting in more severe hyperglycemia and the development of symptoms (Katsarou et al.
-2017).
-Primary symptoms include polyuria, polydipsia, and polyphagia (R A Guthrie and D W Guthrie 2004).
-As glucose accumulates within the bloodstream, the kidneys attempt to
-rectify the imbalance by excreting more glucose in the urine (R A Guthrie and D W Guthrie 2004). As the nephron excretes
-glucose into the urine, water follows from the bloodstream into the urine, resulting in a greater volume of urine and the primary
-symptom of polyuria (R A Guthrie and D W Guthrie 2004). Increased urination in attempts to reduce blood glucose results
-in dehydration, so this causes the polydipsia as the body’s mechanism to increase water intake and restore hydration and
-water homeostasis (R A Guthrie and D W Guthrie 2004). The final primary symptom, polyphagia, is increased hunger due to
-reduced fuel available to cells (R A Guthrie and D W Guthrie 2004). Primary complications include nephropathy, neuropathy,
-retinopathy, risk of cardiovascular disease and peripheral artery disease, and risk of cerebrovascular disease (Katsarou et al.
-2017; McGill and Levitsky 2016; Melendez-Ramirez, Richards, and Cefalu 2010; Wherrett et al. 2018). These conditions
-are largely related to the severity and length of hyperglycemic exposure, especially due to the increased stress on the kidneys
-due to increased glucose excretion. As hyperglycemia exposure increases, this can increase the risk of albuminuria, which
-may progress to complete nephropathy (Melendez-Ramirez, Richards, and Cefalu 2010). Neuropathy may arise in the case of
-hypoglycemia (in instances where too much insulin is injected), as the brain primarily relies on glucose as its fuel source, and
-reduced glucose delivery to the brain may result in neuronal death and possible cognitive impairments and deficits (Katsarou
-et al. 2017). Hyperglycemia is also linked to the production of microaneurysms in blood vessels feeding the retina, leading
-to possible retinopathy and fluid leakage resulting in macular edema (Melendez-Ramirez, Richards, and Cefalu 2010). Lastly,
-hyperglycemia can trigger a severe complication known as diabetic ketoacidosis (DKA). Hyperglycemia causes the production
-of ketones as an alternate source of fuel when insulin is critically low, and cells cannot access glucose (Cashen and Petersen
-2019). In this case, ketogenesis begins, and can result in the acidification of the blood, and can also result in cerebral edema
-and hypokalemia which can become fatal in severe cases (Cashen and Petersen 2019).
+\subsection{Do Games Help or Hurt Learning: The Debate Over Game-Based Learning}
 
-\subsection{Mutations associated with Type 1 Diabetes}
-T1D is a complex disease, meaning it can be influenced by a variety of environmental and genetic risk factors (Brorsson
-et al. 2010). The concordance of T1D among monozygotic and dizygotic twins, as well as numerous studies that have identified
-genes associated with T1D, provide good evidence for a strong genetic determinant of disease (Erlich et al. 2008). For over
-thirty years, it has been known that the human leukocyte antigen (HLA) region on chromosome 6 contributes to the risk of
-developing T1D (Brorsson et al. 2010). The HLA region contains genes which code for proteins that have immune function,
-particularly in the regulation of the immune response (Choo 2007). More specifically, it has been found that certain alleles at
-the HLA-DRB1, HLA-DQA1, and HLA-DQB1 loci are more strongly associated with T1D than others (Erlich et al. 2008).
-Although studies have identified these specific alleles, the consensus remains that more work must be done to further identify
-causative variants and their DNA profile (Brorsson et al. 2010). The complexity of T1D makes it extremely difficult to narrow
-down the cause to one single gene, but there are a very small number of diabetes cases which do in fact result from single
-gene mutations (Yang and Chan 2016). Two major forms of monogenic diabetes are Neonatal diabetes mellitus (NDM) and
-maturity-onset diabetes of the young (MODY) (Yang and Chan 2016). These monogenic forms of diabetes are typically the
-result of mutations in genes which code for proteins that reduce $\beta$-cell function (Steck and Winter 2011). It is interesting to
-note that the single gene mutations found to cause NDM and MODY, have also been found to be associated with T1D (Yang
-and Chan 2016). Two examples of monogenic diabetes genes which have also been found to be associated with T1D are the
-preproinsulin gene (INS), and the Gli-similar 3 gene (GLIS3) (Yang and Chan 2016). The GLIS3 gene codes for a transcription
-factor which regulates islet development and insulin gene transcription, and the INS gene encodes the peptide hormone insulin
-(Steck and Winter 2011; Yang and Chan 2016). Single nucleotide polymorphisms (SNPs) are typically what affect the normal
-function of both genes, but it has also been found that allelic variants of the genes lead to both monogenic diabetes and T1D
-(Yang and Chan 2016).
+The debate of whether game based learning helps or hinders learning is ongoing.
+What comes into question in opposition to game based learning is that it serves to act as a source of entertainment and amusement rather than for educative purposes.
+Many individuals associate play with children, and therefore believe that play is trivial and unimportant (Play and learning Prensky).
+In fact, one does not learn to play, but rather plays to learn.
+Play is where a spark is ignited and the source in which curiosity is peaked.
+It is not silly, trivial or irrelevant but has a biological and evolutionary function which has a strong connection to learning (Play and Learning Presnky).
+It is, as Danny Hillis, founder of Thinking Machines and a former Disney Fellow has stated something every single culture does (play and learning Prensky).
 
-\subsection{Type 1 Diabetes Treatments and Health Considerations}
-As the $\beta$-cell functionality is destroyed due to immune destruction, functionality cannot be restored. Thus, the standard
-of care involves continuous glucose monitoring and regulation of blood sugar levels via insulin injections (McGill and Levitsky
-2016). Many patients opt for insulin pumps, otherwise known as a closed-loop pancreas system (Wherrett et al. 2018). This
-system is an artificial system that features an insulin pump, and a glucose sensor which can detect when insulin should be
-secreted. Some of these systems may infuse just insulin, others also feature a glucagon pump to increase blood sugar in case
-too much insulin was secreted (Wherrett et al. 2018). Alternatively, patients may opt for standard insulin injections. A key
-component of T1D treatment is patient education on their condition and the body’s natural fluctuations of blood sugar levels
-throughout the day. For example, it is important for patients to understand that prolonged fasting, illness, or intense exercise can
-reduce blood sugar levels and naturally cause a hypoglycemic crisis (McGill and Levitsky 2016). In the event of hypoglycemia,
-it is important to detect as soon as possible and consume a small amount of sugar (15 grams typically recommended, available
-in juice, candies, or glucose gels) (Katsarou et al. 2017). Lifestyle changes such as eating snacks prior to bedtime and exercising
-earlier in the day can be beneficial in preventing nocturnal hypoglycemia (McGill and Levitsky 2016). Due to the number of
-potential complications which may arise from T1D, physicians should also monitor patients for other comorbidities including
-other autoimmune conditions, kidney failure, dyslipidemia, retinopathy, and other hormonal disease (Katsarou et al. 2017).
-Further, patients are encouraged to consult dieticians to develop healthy meal plans to regulate blood sugar levels and join
-community support groups (Katsarou et al. 2017). T1D patients are at increased risk of mental health issues such as anxiety,
-depression, and eating disorders due to the perceived lack of autonomy and increased stress load due to their disease (Katsarou
-et al. 2017).
+Arguing the merits behind the development and creation of digital games for use in post-secondary classrooms can appear effortless.
+However, what cannot be overlooked at times is that educators spend a great deal of time planning and executing engaging lessons which can become time consuming and onerous.
+Svinicky and McKeachie (2013) argue that the biggest barrier to the use of games therefore is logistic.
+Although games can be fun and engaging, we know that our classrooms today are comprised of diverse students with varying needs and a one size fits all approach does not work.
+
+\subsection{The Benefits of Using Video Games at the Post-secondary Level}
+
+Classrooms are becoming a place where the student is now involved in the learning process.
+Rules, goals and objectives, outcomes, competition, interaction, and story are what Mark Prensky refers to as the six key structural elements of games (Prensky digital game based learning).
+These powerful factors are what set games apart from other engaging activities.
+Traditional forms of educational media lack what video games provide, a high degree of interactivity (coller, scott, mechanical engineering paper).
+Video games challenge individuals to work within specific parameters and respond to situations that occur in the game (coller, scott, mechanical engineering).
+The way in which individuals respond to challenges presented in video games allows them to think critically, strategize, and work collaboratively in an interactive and engaging manner.
+
+Video games are visually appealing, stimulating, and heighten our senses in many ways.
+We become addicted to the hormone boost our brains receive when we achieve results (Lubomir Tomaska).
+They allow us to learn through doing and creating (video games and education Kurt Squire).
+Like play, upon beginning a video game, one is completely immersed into a whole new world without boundaries (coller, scott, engineering).
+Each player is engaged in a creative process where they are free to discover the limitless possibilities contained within the game.
+In a school setting, we are often taught there is one solution or right answer or way of doing something.
+Video games challenge this by giving players multiple entry points and pathways to consider.
+
+The growing evidence of the effectiveness of video games in the classroom is emerging, and the value of game based learning is evident throughout various disciplines.
+In 2005, Northern Illinois University began teaching an undergraduate course in Numerical Methods using a video game called “NIU-Torcs” (Coller, Scott).
+Students were observed spending twice the average amount of time on coursework than other subject areas (Coller, Scott).
+Other games are being developed for nursing education which focus on developing medication calculation skills.
+This is an area of global concern perhaps due to test anxiety and low mathematical self-efficacy in nursing students (Foss, Mordt Medication calculation).
+Game based learning not only provides students with infinite possibilities for discovery but sustains student interest, builds confidence, and helps to improve cognition and learning outcomes.
+
+\subsection{Concluding the Lit Review - I dont know if I wanna put this here}
+
+The long-term positive effects of Game-Based Learning will continue to be researched and studied.
+Although empirical evidence is slowly emerging but still limited, studies and research to date are proving that Game-Based Learning has the potential to transform post secondary classrooms.
+They are moving from teacher centred to student centred learning hubs which serve to create a learning environment in which students are able to take risks, explore, discover, fail and try again.
 

### 2023-01-09 thesis_template/paper0904.tex
All citations added in bibtex

@@
 
-\newpage
-
-\section{Literature Review/Proposal}
-
-\subsection{Introduction - Maybe move to Abstract, it encompasses everything}
 After many years of trying to adapt to educational innovations of the 21st Century, educators will soon be looking into the not too distant future and considering what modern learning in post-secondary education will look like in the 22nd Century.
@@
 – I don’t know if I should mention the game but I like this line.
+
+\newpage
+
+\section{Literature Review/Proposal}
 
@@
 What comes into question in opposition to game based learning is that it serves to act as a source of entertainment and amusement rather than for educative purposes.
-Many individuals associate play with children, and therefore believe that play is trivial and unimportant (Play and learning Prensky).
+Many individuals associate play with children, and therefore believe that play is trivial and unimportant \citep{prensky2001fun}.
 In fact, one does not learn to play, but rather plays to learn.
 Play is where a spark is ignited and the source in which curiosity is peaked.
-It is not silly, trivial or irrelevant but has a biological and evolutionary function which has a strong connection to learning (Play and Learning Presnky).
-It is, as Danny Hillis, founder of Thinking Machines and a former Disney Fellow has stated something every single culture does (play and learning Prensky).
+It is not silly, trivial or irrelevant but has a biological and evolutionary function which has a strong connection to learning \citep{prensky2001fun}.
+It is, as Danny Hillis, founder of Thinking Machines and a former Disney Fellow has stated something every single culture does \citep{prensky2001fun}.
 
@@
 However, what cannot be overlooked at times is that educators spend a great deal of time planning and executing engaging lessons which can become time consuming and onerous.
-Svinicky and McKeachie (2013) argue that the biggest barrier to the use of games therefore is logistic.
+\citet{mckeachie2013mckeachie} argue that the biggest barrier to the use of games therefore is logistic.
 Although games can be fun and engaging, we know that our classrooms today are comprised of diverse students with varying needs and a one size fits all approach does not work.
@@
 Classrooms are becoming a place where the student is now involved in the learning process.
-Rules, goals and objectives, outcomes, competition, interaction, and story are what Mark Prensky refers to as the six key structural elements of games (Prensky digital game based learning).
+Rules, goals and objectives, outcomes, competition, interaction, and story are what Mark Prensky refers to as the six key structural elements of games \citep{prensky2001fun}.
 These powerful factors are what set games apart from other engaging activities.
-Traditional forms of educational media lack what video games provide, a high degree of interactivity (coller, scott, mechanical engineering paper).
-Video games challenge individuals to work within specific parameters and respond to situations that occur in the game (coller, scott, mechanical engineering).
+Traditional forms of educational media lack what video games provide, a high degree of interactivity \citep{coller2009effectiveness}.
+Video games challenge individuals to work within specific parameters and respond to situations that occur in the game \citep{coller2009effectiveness}.
 The way in which individuals respond to challenges presented in video games allows them to think critically, strategize, and work collaboratively in an interactive and engaging manner.
@@
 Video games are visually appealing, stimulating, and heighten our senses in many ways.
-We become addicted to the hormone boost our brains receive when we achieve results (Lubomir Tomaska).
-They allow us to learn through doing and creating (video games and education Kurt Squire).
-Like play, upon beginning a video game, one is completely immersed into a whole new world without boundaries (coller, scott, engineering).
+We become addicted to the hormone boost our brains receive when we achieve results \citep{tomavska2022let}.
+They allow us to learn through doing and creating.
+Like play, upon beginning a video game, one is completely immersed into a whole new world without boundaries \citep{coller2009effectiveness}.
 Each player is engaged in a creative process where they are free to discover the limitless possibilities contained within the game.
@@
 The growing evidence of the effectiveness of video games in the classroom is emerging, and the value of game based learning is evident throughout various disciplines.
-In 2005, Northern Illinois University began teaching an undergraduate course in Numerical Methods using a video game called “NIU-Torcs” (Coller, Scott).
-Students were observed spending twice the average amount of time on coursework than other subject areas (Coller, Scott).
+In 2005, Northern Illinois University began teaching an undergraduate course in Numerical Methods using a video game called “NIU-Torcs” \citep{coller2009effectiveness}.
+Students were observed spending twice the average amount of time on coursework than other subject areas \citep{coller2009effectiveness}.
 Other games are being developed for nursing education which focus on developing medication calculation skills.
-This is an area of global concern perhaps due to test anxiety and low mathematical self-efficacy in nursing students (Foss, Mordt Medication calculation).
+This is an area of global concern perhaps due to test anxiety and low mathematical self-efficacy in nursing students \citep{foss2013medication}.
 Game based learning not only provides students with infinite possibilities for discovery but sustains student interest, builds confidence, and helps to improve cognition and learning outcomes.
@@
 They are moving from teacher centred to student centred learning hubs which serve to create a learning environment in which students are able to take risks, explore, discover, fail and try again.
+
+NOW WE ADD CELLS AT WAR DATA!
 

### 2023-03-06 thesis_template/paper0904.tex
Mar 5 updates - fixed rosas mid-year report edits, started writing materials and methods

@@
 The traditional construct of the classroom has started to evolve, embracing a new way in which knowledge, skills and attitudes can be acquired.
-Cells at War is a collection of scientific video games developed through interdisciplinary efforts that reinforces the notion that video games are highly effective tools used to enhance instructional methods.
-(Not only enhancing but teaching and engaging, ask rosa).
-– I don’t know if I should mention the game but I like this line.
+Cells at War is a collection of scientific video games developed through interdisciplinary efforts that reinforces the notion that video games are highly effective tools used to enhance instructional methods and engage students in the Biology classroom.
 
@@
 
-\subsection{Children and Play}
+\subsection{Gamification of Learning}
 
-Although technological ecosystems have been established in schools for some time, its landscape continues to be missing significant components to create a fully functioning ecosystem.
-One of these key components is that educators are reticent to embrace digital technology and gaming as the new norm in the classroom.
+Although the use of technology has become prevalent in schools across Canada, using game-based learning as an instructional approach is still in its infant stages.
+Educators are reticent to embrace digital technology and gaming as the new norm in the classroom due to insufficient technical competence, as well as challenges they face in selecting appropriate games \citep{jaaska2022teachers,molin2017role}.
+In addition, educators are rarely involved in the development of these games, making it difficult to facilitate suitable learning in the classroom \citep{molin2017role}.
+
 Kindergarten classrooms in Ontario are bustling places of activity, reflective of active inquiry through play.
 Exploration and provocations drive their curiosity while simultaneously developing their oral language and critical thinking skills \citep{kindergartencurriculum}.
-Digital games on ipads are being used to develop literacy and numeracy skills.
-Some educators however, oppose the use of digitized games especially in higher education and have not fully embraced the notion that learning can actually take place through play.
 We believe that play is paramount for children in their formative years and educational theorists like Vygotsky affirm this.
 He was firm in his belief that play is the leading line of development during preschool years \citep{vygotsky1967play}.
-Unfortunately, in many classrooms, educators continue to feel more comfortable in engaging in antiquated and traditional methods of instruction that consist of a didactic approach to learning.
-The need to challenge current educational practices is paramount as educators are currently teaching students from the perspective of their own past rather than the realities of the students’ future.
-– I like this line, its powerful that’s why I highlighted it.
+Unfortunately, in many classrooms, educators continue to feel more comfortable engaging in antiquated and traditional methods of instruction that consist of a didactic approach to learning.
+Some educators may be fearful that incorporating game-based learning into the curriculum will undermine their authority in the classroom \citep{jaaska2022teachers,chee2014facilitating,jong2016teachers}.
+The need to challenge current educational practices is paramount as it appears some educators are currently teaching students from the perspective of their own past rather than the realities of the students’ future.
 
@@
 
-It can be argued that this fear of play is also due in large part to limited resources and training being provided to educators.
+It can be argued that this fear of play is also due in large part to limited resources and training being provided to educators \citep{fishman2014empowering}.
 Teachers require technological and pedagogical supports to develop their understanding of game based learning in the classroom.
@@
 Incorporating game based learning in teacher practice is impacted by a lack of professional development opportunities that can systematically guide them in using games for teaching, learning, and assessment \citep{fishman2014empowering,ruggiero2013video,foster2020principles}.
-Might want to add a line here, don’t wanna end on quote
+In order to develop teacher proficiency in game-based learning, further professional development is required in the form of pedagogical training in order to understand how to develop and utilize game-based learning practices.
 
@@
 The debate of whether game based learning helps or hinders learning is ongoing.
-What comes into question in opposition to game based learning is that it serves to act as a source of entertainment and amusement rather than for educative purposes.
+What comes into question in opposition to game based learning is that it serves to act as a source of entertainment and amusement rather than for educative purposes \citep{lister2015gamification, mayer2010adding}.
 Many individuals associate play with children, and therefore believe that play is trivial and unimportant \citep{prensky2001fun}.
-In fact, one does not learn to play, but rather plays to learn.
+It appears one does not learn to play, but rather plays to learn.
 Play is where a spark is ignited and the source in which curiosity is peaked.
@@
 
-Classrooms are becoming a place where the student is now involved in the learning process.
  Rules, goals and objectives, outcomes, competition, interaction, and story are what Mark Prensky refers to as the six key structural elements of games \citep{prensky2001fun}.
@@
 This is an area of global concern perhaps due to test anxiety and low mathematical self-efficacy in nursing students \citep{foss2013medication}.
-Game based learning not only provides students with infinite possibilities for discovery but sustains student interest, builds confidence, and helps to improve cognition and learning outcomes.
-
-\subsection{Concluding the Lit Review - I dont know if I wanna put this here}
+Further, as a result of the COVID-19 pandemic, educators were compelled to find alternative ways to supplement hands-on learning experiences in an online setting.
+In a study completed at McMaster University, virtual reality (VR) headsets were used to engage in a 3D laboratory simulation which allowed for open-ended investigations in a range of STEM fields such as chemistry, biology, and physics \citep{tsirulnikov2023game}.
+\citet{tsirulnikov2023game} noted higher levels of engagement and motivation through these VR simulations.
+Game based learning not only provides students with infinite possibilities for discovery but sustains student interest, builds confidence, and helps to improve cognition and learning outcomes \citep{divjak2011impact}.
 
@@
 Although empirical evidence is slowly emerging but still limited, studies and research to date are proving that Game-Based Learning has the potential to transform post secondary classrooms.
-They are moving from teacher centred to student centred learning hubs which serve to create a learning environment in which students are able to take risks, explore, discover, fail and try again.
+A learning environment in which students are able to take risks, explore, discover, fail and try again is achieved through a collaborative and interactive process between student and teacher.
 
-NOW WE ADD CELLS AT WAR DATA!
+\section{Materials and Methods}
+
+\subsection{Study Setting and Context}
+
+Results from this study were obtained utilizing a first-year cellular and molecular biology course (BIO1A03) at McMaster University (Ontario, Canada).
+BIO1A03 provides first year students with a strong cell and molecular biology foundation that is required for most upper level courses in the Faculty of Science at McMaster.
+Enrollment in BIO1A03 tends to be high, with approximately 700 students taking the course during both the fall (September-December) and winter (January-April) semesters.
+Due to the lengthy process of video game design, we were limited to collecting data solely from students enrolled in the winter semester for this study.
+
+It is worthwhile to note the current course format of BIO1A03.
+Prior to 2014, BIO1A03 followed a traditional, didactic format, which consisted of three 50 minute in-person lectures per week, accompanied by a bi-weekly, three-hour laboratory session \citep{tahir2022blended}.
+In 2013, BIO1A03 was converted into a Blended Learning course format, which included independent, self-paced learning through prerecorded modules, and face-to-face learning in the classroom with a heavier focus on student collaboration and engagement \citep{tahir2022blended}.
+Course content previously delivered during lecture was condensed into two weekly modules which students were required to complete prior to attending in-person lectures.
+In-person lectures were restructured and minimized to two per week.
+One of these sessions consisted of a review lecture, where students could clarify concepts they found difficult to understand, while the other focused on applying theoretical concepts that students had learned to a real world concept.
+In a study highlighting the impacts of blended learning in BIO1A03 at McMaster, \citet{tahir2022blended} noted that the majority of students surveyed after taking the course felt that these novel blended learning approaches enabled them to improve general knowledge and understanding of core concepts in cellular and molecular biology.
+
+\subsection{Video Game Design/Creation}
+
+The conversion to a blended learning approach in BIO1A03 allowed for a seamless transition to online-learning in 2019 when the COVID-19 pandemic began.
+During this time, the BIO1A03 instructional team began exploring other novel pedagogical approaches which could help improve student learning.
+The idea of using digital games to improve education has been around for quite some time, however the controversy surrounding video games in this context is still apparent.
+
+In 2021, the BIO1A03 instructional team partnered up with the George Brown College School of Design, and received support from CEWIL Canada (Co-operative Education and Work-Integrated Learning Canada), formerly known as CAFCE (Canadian Association for Co-operative Education), to create and develop an educational biological video game that could be used in BIO1A03.
+
+\subsection{Study Participants}
+
+\subsection{Survey Design: Qualitative Analysis of Student Feedback}
+
+\section{Results}
+
+\section{Discussion}
+
+\section{Future Directions}
 

### 2023-03-07 thesis_template/paper0904.tex
March 7, basically done materials and methods

@@
 
-In 2021, the BIO1A03 instructional team partnered up with the George Brown College School of Design, and received support from CEWIL Canada (Co-operative Education and Work-Integrated Learning Canada), formerly known as CAFCE (Canadian Association for Co-operative Education), to create and develop an educational biological video game that could be used in BIO1A03.
+In 2021, the BIO1A03 instructional team partnered up with the George Brown College School of Design, and received support from CEWIL Canada (Co-operative Education and Work-Integrated Learning Canada), formerly known as CAFCE (Canadian Association for Co-operative Education), to design and implement an educational biological video game that could be used as an instructional and engaging teaching tool in BIO1A03.
+A small cohort of McMaster Univeristy Biology students and supervising professors, as well as game designers, programmers, animators, audio engineers, and artists from the George Brown College School of Design were selected to pilot this project.
+
+Concept development began in September of 2021.
+In this phase, students from both institutions worked collaboratively to brainstorm concepts for a biological video game that would be both engaging and educational.
+The theme for the game was centred around single-gene diseases.
+Biology students from McMaster University were responsible for selecting and researching a specific single-gene disease.
+This information was exchanged with students from George Brown College, who in-turn provided a wealth of knowledge on how to turn these static, difficult to understand concepts into engaging gameplay mechanics.
+Students pitched these proposals to supervising professors who then selected the ideas and game elements that best suited the goals of the project.
+
+The game production phase then began in November of 2021.
+Students from George Brown College started bringing these ideas to life through artwork, animation, and programming.
+McMaster Biology students served as subject matter experts, ensuring scientific accuracy throughout the production process, as well as creating an encyclopedia of important biological concepts that were highlighted in-game.
+Throughout the production stage, weekly online meetings were held to keep all members updated with progress.
+Discord was also utilized throughout the entire process in order to ensure open dialogue between students and supervisors, as well as to share game assets and important information.
+A functioning build of the game was completed in late March of 2021, and was introduced in the final BIO1A03 class of the winter 2021 semester, in place of lecture material.
+After class, students were asked to complete a survey regarding their attitudes towards the game.
 
@@
 
+Students qualifying for participation in the survey included only those enrolled in the winter 2021 semester of BIO1A03.
+Students were informed about the survey prior to playing the game, through both in-class and online announcements posted on Desire to Learn.
+Students who did not complete the survey were informed that this would not effect course performance.
+Those who participated were incentivized by being entered into a draw to win prizes.
+
 \subsection{Survey Design: Qualitative Analysis of Student Feedback}
+
+The survey was developed and delivered through the LimeSurvey\textregistered \space platform.
+Students were asked a total of 14 questions regarding opinions and attitudes on their current undergraduate education, as well as feedback on game-based learning in the biology classroom.
+The National Survey of Student Engagement (NSSE) was used as a framework to generate the questions that were asked.
+NSSE questions assess the extent to which students engage in educational practices in the context of higher levels of learning (NSSE website citation).
+MORE HERE ABOUT NSSE...
 

### 2023-03-08 survey_stuff_2022_cellsatwar_foranalysis/survey_analysis.R
added survey analysis to github

@@ -1,0 +1,276 @@
+###CELLS AT WAR 2022 SURVEY DATA ANALYSIS - BEGAN FEB 23
+
+###IMPORTING THE DATA
+morning_df_csv <- read.csv("Depersonalized_survey_data_morning(2).csv", header = TRUE)
+combined_df_csv <- read.csv("combined_data_morning_afternoon_raw_CSV.csv", header = TRUE)
+
+###CREATING SOME GRAPHS
+library(ggplot2)
+library(ggthemes)
+library(ggpubr) #For ggarrange function
+
+### FUNCTION FOR PUMPING OUT BARPLOTS
+plot_survey_fun <- function(df, bar_lables, xtitle, ytitle, chart_title) {
+  p <- ggplot(df, aes(x = df[,1], y = df[,2])) + 
+    geom_col(show.legend = FALSE, width = 0.7, fill = "#000075") +
+    geom_text(aes(label = df[,2]), vjust = -0.5, colour = "black", nudge_x = -0.14 ) +
+    geom_text(aes(label = paste0("(",round(df[,2]/sum(df[,2])*100),")%")),vjust = -0.5, nudge_x = 0.14 ) +
+    scale_x_discrete(labels = bar_lables) +
+    xlab(xtitle) +
+    ylab(ytitle) +
+    ggtitle(chart_title)
+  
+  p + theme_clean(base_size = 8)
+}
+
+#Q1
+########################################################################
+###ON AVERAGE, HOW MUCH TIME DO YOU SPEND PLAYING VIDEO GAMES EACH WEEK#
+########################################################################
+
+time_spent_vg <- as.data.frame(table(combined_df_csv$On.average..how.much.time.do.you.spend.playing.video.games.each.week....Select.one.option.))
+time_spent_vg_labels <- c("NA", "I don't play video games", "Between 10-20 hours/week", "3-10 hours/week", "Under 3 hours a week", "More than 20 hours/week")
+time_spent_vg_xtitle <- "Responses"
+time_spent_vg_ytitle <- "Frequency"
+time_spent_vg_title <- "Time Spent Playing Video Games"
+
+##SAVING THE PLOT TO DIRECTORY
+tiff("figures/time_spent_playing_videogames.tiff", units="in", width=8, height=6, res=300)
+plot_survey_fun(time_spent_vg, time_spent_vg_labels, time_spent_vg_xtitle, time_spent_vg_ytitle, time_spent_vg_title)
+dev.off()
+
[…]

### 2023-03-08 cellsatwar_figures/beap_pres_dec.tex
ppt of figures created march 8

% Inbuilt themes in beamer
\documentclass{beamer}

%Defining some colours
\definecolor{darkred}{rgb}{0.8,0,0}

% Theme choice: there a number of preset themes to choose from
% Play around with them, Cambridge is nice for first pres
%\usetheme{Szeged}
\usetheme{CambridgeUS} %setting the main theme
%\usecolortheme{beaver} % setting colour theme
\usefonttheme{professionalfonts} %font theme
\useinnertheme[shadow=true]{rounded}
%\useoutertheme{} %outer theme

%\setbeamertemplate{footline} %Remove footer line in all slides
\setbeamertemplate{navigation symbols}{} %removes navigation symbols
\setbeamertemplate{footline}[page number] %removes footer line, keeps pg#
\setbeamertemplate{caption}{\insertcaption}

%Setting colours for boxes and captions
\setbeamercolor{block title}{bg=darkred, fg=white}
\setbeamercolor{block body}{bg=darkred!10}

% For the Flowchart
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

% Title page details: 
\title[BEAP Dec 2022]{Cells at War 2021 Survey Data } 
\author{Alexander Turco}
\date{March 10, 2022}
%\logo{\includegraphics[height=1cm, width=1cm]{logo.png}}

% Bibliography stuff
\usepackage[natbib=true, sorting=nyt, style=authoryear-comp]{biblatex}
\addbibresource{ABC1.bib}

%Extra packages
\usepackage{makecell}

%For itemize
\setbeamertemplate{itemize item}[triangle]

\begin{document}

	% For my introduction slides, there will be the slide with my title and name as well as an outline slide with the brief overview of what I will discuss.
	% Title page frame - SLIDE 1%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\section{Figures}
	\begin{frame}
		\titlepage 
	\end{frame}

	% Remove logo from the next slides
	%\logo{}

	% Outline frame - SLIDE 2
	\begin{frame}{On average, how much time do you spend playing video games each week?}

		\begin{center}
			\includegraphics[width=10cm, height=7cm]{figures/time_spent_playing_videogames.jpg}
		\end{center}

	\end{frame}

	\begin{frame}{What types of courses have you taken at McMaster this current school year?}

		\begin{center}
			\includegraphics[width=10cm, height=7cm]{figures/types_of_courses_taken.jpg}
		\end{center}

	\end{frame}

	\begin{frame}{During the current school year, how often would you say you have done the following?}

		\begin{center}
			\includegraphics[width=10cm, height=7cm]{figures/how_often_haveyou_done_thefollowing.jpg}
		\end{center}

	\end{frame}

	\begin{frame}{During the current school year, how much would you say your coursework has emphasized the following? }

		\begin{center}
			\includegraphics[width=10cm, height=7cm]{figures/howmuch_coursework_emphasized_thefollowing.jpg}
		\end{center}

	\end{frame}

	\begin{frame}{If game-based learning approaches were incorporated into one of your science courses, how do you think that this mode of learning can help with the following? }

		\begin{center}
			\includegraphics[width=10cm, height=6cm]{figures/how_does_gbl_help_in_science.jpg}
		\end{center}

	\end{frame}

	\begin{frame}{If BIO1A03 added a game-based learning component, how much would this improve your motivation to do the following? }

		\begin{center}
			\includegraphics[width=10cm, height=6cm]{figures/ifbio1a03_added_gblcomponent.jpg}
		\end{center}

	\end{frame}

	\begin{frame}{How likely is it that you would play the video game Cells at War on your own time, outside of class to further consolidate material taught during this class? }

		\begin{center}
			\includegraphics[width=9cm, height=6cm]{figures/how_likely_you_would_play_cellsatwar.jpg}
		\end{center}

	\end{frame}

	\begin{frame}{How prepared would you feel if you were given a quiz on Pompe Disease based on the Cells at War game, compared to studying off traditional lecture slides (with accompanying readings)? }

		\begin{center}
			\includegraphics[width=9cm, height=6cm]{figures/how_prepared_if_tested_on_pompe_based_off_cellsatwar.jpg}
		\end{center}

	\end{frame}

\end{document}

### 2023-03-13 thesis_template/paper0904.tex
Updated Abstract

@@
 
-After many years of trying to adapt to educational innovations of the 21st Century, educators will soon be looking into the not too distant future and considering what modern learning in post-secondary education will look like in the 22nd Century.
-Technological advances will continue to transform the landscape of the classroom and the use of digital games, although a source of controversy and debate, will prove to be highly effective instructional tools that can promote problem solving skills and critical thinking.
-Their use has already started to change the structure of the class in the elementary classroom as they serve to energize and engage students in their learning and improve learning, creativity and imagination.
-The traditional construct of the classroom has started to evolve, embracing a new way in which knowledge, skills and attitudes can be acquired.
-Cells at War is a collection of scientific video games developed through interdisciplinary efforts that reinforces the notion that video games are highly effective tools used to enhance instructional methods and engage students in the Biology classroom.
+After many years of trying to adapt to educational innovations of the 21st Century, educators will soon be looking into the not-too-distant future to consider what modern learning in post-secondary education will look like in the 22nd Century.
+Technological advances will continue to transform the landscape of the classroom.
+The use of digital games, although a source of controversy and debate, will prove to be highly effective instructional tools that will promote problem solving and critical thinking skills.
+Their use has already started to change the structure of the class in the elementary classroom as they serve to energize and engage students in their learning and improve conceptual understandings, creativity, and imagination.
+The traditional construct of the classroom has started to evolve, embracing a new way in which knowledge, skills, and attitudes can be acquired.
+The primary focus of this study was to create and implement a biological video game that focused on a single gene disease, which served as a vehicle to teach cellular and molecular biology concepts.
+‘Cells at War’ is a collection of scientific video games developed through interdisciplinary efforts that reinforces the notion that video games are highly effective tools used to enhance instructional methods and engage students in the biology classroom.
 

### 2023-03-14 thesis_template/paper0904.tex
Results inputted into paper, start writing now

@@
 \DeclareUnicodeCharacter{0301}{\'{e}}
+\usepackage{float}
 \usepackage{times}
@@
 \title{\sc Cells at War: The Playfulness of Game-Based Learning}
-\author{\sc Bianca Flaim, Alexander Turco and Rosa da Silva}
+\author{\sc Alexander Turco and Rosa da Silva}
 
@@
 In 2021, the BIO1A03 instructional team partnered up with the George Brown College School of Design, and received support from CEWIL Canada (Co-operative Education and Work-Integrated Learning Canada), formerly known as CAFCE (Canadian Association for Co-operative Education), to design and implement an educational biological video game that could be used as an instructional and engaging teaching tool in BIO1A03.
-A small cohort of McMaster Univeristy Biology students and supervising professors, as well as game designers, programmers, animators, audio engineers, and artists from the George Brown College School of Design were selected to pilot this project.
+A small cohort of McMaster Univeristy Biology students and supervising professors, as well as game designers, programmers, animators, audio engineers, and artists from the George Brown College School of Design were selected to pilot this work-integrated learning project.
 
@@
 
-\subsection{Survey Design: Qualitative Analysis of Student Feedback}
+\subsection{Survey Design: Qualitative Analysis of Student Feedback NOT DONE YET}
 
@@
 
+\subsection{Student perceptions of current undergraduate experiences}
+
+150 students enrolled in the winter 2021 session of BIO1A03 attended a game-based style lecture and completed the follow-up survey.
+Participants were asked a series of questions regarding their current experiences as undergraduate science students, as well as questions regarding the benefits of implementing game-based learning approaches into their classrooms.
+
+%%%%%TIME SPENT PLAYING VIDEO GAMES
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{figures_4f06/time_spent_playing_videogames.jpg}
+	\caption{Student responses regarding the average amount of weekly time spent playing video games}
+	\label{fig:1}
+\end{figure}
+
+%%%%%Types of courses taken by students
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{figures_4f06/types_of_courses_taken.jpg}
+	\caption{Types of Courses Taken}
+	\label{fig:2}
+\end{figure}
+
+%%%%%During the current school year how much has your coursework emphasized the following
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{figures_4f06/howmuch_coursework_emphasized_thefollowing.jpg}
+	\caption{How Much Has Coursework Emphasized the following}
+	\label{fig:3}
+\end{figure}
+
+%%%%%During the current school year how often would you say you have done the following
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{figures_4f06/how_often_haveyou_done_thefollowing.jpg}
+	\caption{How often have you done the following}
+	\label{fig:4}
+\end{figure}
+
+\subsection{Student perceptions on the use of game-based learning in biology}
+
+%%%%%If game-based learning approaches were incorporated into one of your science courses, how do you think that this mode of learning can help with the following
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{figures_4f06/how_does_gbl_help_in_science.jpg}
+	\caption{If game-based learning approaches were incorporated in your science courses}
+	\label{fig:5}
+\end{figure}
+
+%%%%% If BIO1A03 added a game-based learning component, how much would this improve your motivation to do the following?
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{figures_4f06/ifbio1a03_added_gblcomponent.jpg}
+	\caption{If BIO1A03 added a game-based component}
+	\label{fig:6}
+\end{figure}
+
+%%%%% How likely is it that you would play the video game Cells at War on your own time, outside of class to further consolidate material taught during this class
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{figures_4f06/how_likely_you_would_play_cellsatwar.jpg}
+	\caption{How likely is it you would play Cells at War outside of class}
+	\label{fig:7}
+\end{figure}
+
+%%%%%How prepared would you feel if you were given a quiz on Pompe Disease based on the Cells at War game, compared to studying off traditional lecture slides
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{figures_4f06/how_prepared_if_tested_on_pompe_based_off_cellsatwar.jpg}
+	\caption{How prepared would you feel if given a quiz on Pompe Disease}
+	\label{fig:8}
+\end{figure}
+
 \section{Discussion}
+
+We need to highlight in this section the use of Cells at War as an instructional tool as well as the benefits that students get out of this work integrated learning experience
 

### 2023-03-15 thesis_template/paper0904.tex
Results writingcoming together

@@
 \lhead{Cells at War: Game-Based Learning} %This needs to change
-\rhead{Bianca Flaim, Alexander Turco}
+\rhead{Alexander Turco and Rosa da Silva}
 \title{\sc Cells at War: The Playfulness of Game-Based Learning}
@@
 150 students enrolled in the winter 2021 session of BIO1A03 attended a game-based style lecture and completed the follow-up survey.
+Students were presented with condensed lecture material prior to completing the relevant level in the game.
+The idea behind this was to provide a brief introduction of the content to students, and then allow them to further explore these concepts by playing the game.
+
 Participants were asked a series of questions regarding their current experiences as undergraduate science students, as well as questions regarding the benefits of implementing game-based learning approaches into their classrooms.
+To gain insight into how experience students had playing video games, they were asked to indicate a weekly average amount of time spent playing them \figref{fig:1}.
+66 Participants (44\%) indicated that they did not play video games, while 53 participants (36\%) indicated that they spent between 1-10 hours per week.
+Only 7 students (5\%) were found spending approximately 10-20 hours per week playing video games.
+It is interesting to note that participants who reported not playing video games also emphasized that the game was `easy to navigate' and `very accessible'.
+This suggests that a lack of experience playing video games did not necessarily limit students' abilities to learn course content through playing the game.
 
@@
 \end{figure}
+
+Further data obtained provided insight into the various types of courses students were taking at McMaster University during the fall/winter 2021 semesters \figref{fig:2}.
+At this time, McMaster University was slowly shifting back to in-person classes due to COVID-19 restrictions being lifted, however, a large majority of courses had been converted to hybrid, blended learning, and online courses.
+60 participants (40\%) indicated that the majority of courses they were taking were mostly hybrid or blended learning style courses.
+28 participants (19\%) indicated most of their courses were back in-person, while only 7 participants (5\%) were enrolled in mostly online courses.
+30 participants indicated a balanced mix of all course types.
+These results highlight the shift from traditional teaching practices towards newer, more innovative educational practices.
 

### 2023-03-19 thesis_template/paper0904.tex
another results push
Minor updates to results

@@
 30 participants indicated a balanced mix of all course types.
-These results highlight the shift from traditional teaching practices towards newer, more innovative educational practices.
+These results highlight the shift from traditional teaching practices towards more innovative educational practices.
 
@@
 \end{figure}
+
+Students were asked to reflect on their experiences as undergraduate students at McMaster University.
+Participants responded to two questions regarding how much coursework has emphasized certain aspects of their learning \figref{fig:3}, and how often they have applied their learning in different contexts \figref{fig:4}.
+62
 

### 2023-03-28 cellsatwar_bus_presentation/beap_pres_dec.tex
BUS presentation and script DONE - practice now

% Inbuilt themes in beamer
\documentclass{beamer}

%Defining some colours
\definecolor{darkred}{rgb}{0.8,0,0}

% Theme choice: there a number of preset themes to choose from
% Play around with them, Cambridge is nice for first pres
%\usetheme{Szeged}
\usetheme{CambridgeUS} %setting the main theme
\usecolortheme{whale} % setting colour theme
\usefonttheme{professionalfonts} %font theme
\useinnertheme[shadow=true]{rounded}
%\useoutertheme{} %outer theme

%\setbeamertemplate{footline} %Remove footer line in all slides
\setbeamertemplate{navigation symbols}{} %removes navigation symbols
\setbeamertemplate{footline}[page number] %removes footer line, keeps pg#
\setbeamertemplate{caption}{\insertcaption}

%Setting colours for boxes and captions
\setbeamercolor{block title}{bg=blue!30, fg=black}
\setbeamercolor{block body}{bg=blue!10}
\setbeamercolor{frametitle}{fg=black}

%%%FOR VIDEOS
\usepackage{media9}
\usepackage{multimedia}
% For the Flowchart
%\usepackage{tikz}
%\usetikzlibrary{shapes.geometric, arrows}
%\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
%\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
%\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
%\tikzstyle{arrow} = [thick,->,>=stealth]

%%% For the Timeline
\usepackage{xcolor}
\usepackage{tikz} \usetikzlibrary{calc, arrows.meta, intersections, patterns, positioning, shapes.misc, fadings, through,decorations.pathreplacing}

\definecolor{ColorOne}{rgb}{0.0,0.5,1.0} %Lightblue
\definecolor{ColorTwo}{rgb}{1.0,0.6,0.4} %lightorange
\definecolor{ColorThree}{rgb}{0.75,0.58,0.89} %lightpurple

% Title page details: 
\title[BEAP Dec 2022]{Cells at War: The Playfulness of Game-Based Learning} 
\author{Alexander Turco}
\date{December 5, 2022}
\logo{\includegraphics[height=1cm, width=1cm]{logo.png}}

% Bibliography stuff
\usepackage[natbib=true, sorting=nyt, style=authoryear-comp]{biblatex}
\addbibresource{ABC1.bib}

%Extra packages
\usepackage{makecell}

%For itemize
\setbeamertemplate{itemize item}[triangle]

\begin{document}

	% For my introduction slides, there will be the slide with my title and name as well as an outline slide with the brief overview of what I will discuss.
	% Title page frame - SLIDE 1%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\section{Introduction}
	\begin{frame}
		\titlepage 
		\begin{center}
			\includegraphics[width=7cm, height=4cm]{cellsatwar.png}
		\end{center}
	\end{frame}

	% Remove logo from the next slides
	\logo{}

	% Outline frame - SLIDE 2
	\begin{frame}{Overview}

		\begin{center}
		\begin{minipage}{6cm}

		  		\begin{block}{} \hyperlink{link1}{Background Information} \end{block}
		  		\begin{block}{} The Game Design Process \end{block}
		  		\begin{block}{} Collection of Student Feedback \end{block}
		  		\begin{block}{} Results \end{block}
		  		\begin{block}{} Conclusions \end{block}
		  		\begin{block}{} Future Work \end{block}

		\end{minipage}
		\end{center}

	\end{frame}

	% For my Background slides, I will talk about important things such as WHat LCRs are, their evolution, stuff like that
	% SLIDE 3 - WHAT ARE LCRs%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% The graphic for this slide is an image of some output from segA with the parameters in the brackets
	\section{Background}
	\begin{frame}{What is Game-based Learning?}
		\label{link1}

		\begin{itemize}
			\item Utilizing games in educational settings as an interactive medium to promote {\color{blue}\textbf{higher-order thinking}}, {\color{blue}\textbf{social skills}}, and {\color{blue}\textbf{problem-solving skills}} \newline
			\item Games support learning through {\color{blue}\textbf{engagement}}, {\color{blue}\textbf{motivation}}, {\color{blue}\textbf{interactivity}}, 
			{\color{blue}\textbf{drill and practice}}, and {\color{blue}\textbf{content mastery}} \newline
			\item In the digital era, game-based learning commonly refers to the use of {\color{blue}\textbf{digital games}} as effective instructional tools which also help develop {\color{blue}\textbf{digital proficiency}}

			\footnotetext[1]{\tiny\cite{brown1989situated, jan2016game}}
		\end{itemize}

	\end{frame}

	% SLIDE 4 - I do Not Know if I am including this slide
	% The Graphic for this slide is an  
	% Entropy, which is measured by Shannon's Entropy equation is a measure of compositional complexity which uses the proportion of residue(s)
	% in a subsequence to measure the compositional state of that subseequence - A lower variety of residues = lower entropy 
	%\begin{frame}{Shannon's Entropy - MAYBE }

	%		$H = -L\sum p_i log_2(p_i)$

	%\end{frame}

	%SLIDE 5 - LCR's PRESENT IN UNIQUE WAYS
	\begin{frame}{The Benefits of Playing Video Games}

		\begin{block}{\textit{Interactive}}
			 \begin{itemize}
			 \item Individuals respond to situations presented in games, allowing them to think critically and strategize
			 \end{itemize}
		\end{block} \pause

		\begin{block}{\textit{Stimulating}}
			\begin{itemize}
				\item Individuals become addicted to the hormone boost their brains receive when they achieve results
			\end{itemize}
		\end{block} \pause

		\begin{block}{\textit{Exploratory}}
			\begin{itemize}
				\item Individuals are immersed in a world without boundaries, free to explore in any way they wish
			\end{itemize}
		\end{block}

	\footnotetext[2]{\tiny\cite{coller2009effectiveness, tomavska2022let}}

	\end{frame}

	%SLIDE 6 Limitations of Implementing Game-Based Learning
	\begin{frame}{The Limitations of Implementing Game-based Learning}
		\begin{enumerate}
			\item Many individuals believe that playing video games in a class setting will serve to act only as a source of entertainment and amusement rather than for educative purposes \newline 
			\item Educators are rarely involved in the development of educational video games \newline
			\item Teachers lack the technological and pedagogical supports to develop their understanding of game-based learning in the classroom

		\end{enumerate}

	\footnotetext[3]{\tiny\cite{lister2015gamification,mayer2010adding,molin2017role,fishman2014empowering}}

	\end{frame}

	%%%%% APPRACH SECTION BEGINS HERE, 2 SLIDES
	\section{Approach}

	\begin{frame}{Cells at War Game Design Process}
		\tikzstyle{descript} = [text = black,align=center, minimum height=1.8cm, align=center, outer sep=0pt,font = \footnotesize]
		\tikzstyle{activity} =[align=center,outer sep=1pt]

		\begin{tikzpicture}[very thick, black]
			\small

			%% Coordinates
			\coordinate (O) at (-0.25,0); % Origin
			\coordinate (P1) at (4,0);
			\coordinate (CD) at (3,0);
			\coordinate (GP) at (9,0);
			\coordinate (P2) at (8,0);
			\coordinate (P3) at (11,0);
			\coordinate (E1) at (10,0);
			\coordinate (D1) at (1.7,0);
			\coordinate (F) at (11.5,0); %End

			%% Filled regions
			\fill[color=ColorOne!20] rectangle (O) -- (CD) -- ($(CD)+(0,1)$) -- ($(O)+(0,1)$); % Studies
			\fill[color=ColorTwo!20] rectangle (CD) -- (GP) -- ($(GP)+(0,1)$) -- ($(CD)+(0,1)$); % Work

			%% Text inside filled regions
			\draw ($(P1)+(-2.5,0.5)$) node[activity,ColorOne] {Brainstorm};
			\draw ($(P2)+(-2,0.5)$) node[activity,ColorTwo] {Game Production};

			%% Description
			\node[descript,fill=ColorTwo!15,text=ColorTwo](D2) at ($(P2)+(-2,-2.5)$) {%
				\textbf{Where}\\
				assets are\\
				produced};

			\node[descript,fill=ColorOne!15,text=ColorOne](D3) at ($(D1)+(1,-2.5)$) {%
				\textbf{Where}\\
				concept development\\
				begins};

			%% Events
			\draw[<-,thick,color=black] ($(E1)+(0,0.1)$) -- ($(E1)+(0,1.5)$) node [above=0pt,align=center,black] {Game Launch\\BIO1A03};

			%% Arrows
			\path[->,color=ColorTwo] ($(P2)+(-1,-0.1)$) edge [out=-90, in=130]  ($(D2)+(0,1)$);
			\path[->,color=ColorOne]($(D1)+(-1,-0.1)$)  edge [out=-70, in=90]  ($(D3)+(0,1)$);

			%% Arrow
			\draw[->] (O) -- (F);
			%% Ticks
			\foreach \x in {0,1.5,...,12}
			\draw(\x cm,3pt) -- (\x cm,-3pt);
			%% Labels
			\foreach \i \j in {0/Sept,1.5/Oct,3/Nov,4.5/Dec,6/Jan, 7.5/Feb, 9/Mar, 10.5/Apr}{
				\draw (\i,0) node[below=3pt] {\j} ;
			}

		\end{tikzpicture}

	\end{frame}

	\begin{frame}{Example Assets}
		\begin{columns}
			\begin{column}{0.33\textwidth}
				\includegraphics[width=4cm, height=3cm]{asset1.png}
				\includegraphics[width=4cm, height=3cm]{asset2.png}
			\end{column}
			\begin{column}{0.33\textwidth}
				\includegraphics[width=4cm, height=3cm]{asset3.png}
				\includegraphics[width=4cm, height=3cm]{asset4.png}
			\end{column}
			\begin{column}{0.33\textwidth}
				\includegraphics[width=4cm, height=3cm]{asset5.png}
				\includegraphics[width=4cm, height=3cm]{asset6.png}
			\end{column}
		\end{columns}		
	\end{frame}

	\begin{frame}{Some Gameplay}
		\centering
		\movie[width=10cm,height=7cm,showcontrols=true]{\includegraphics[width=10cm, height=7cm]{pompe_thumbnail.png}}{pompe_trimmed.mp4}
	\end{frame}

	%%% EXPLAIN NSSE SURVEY A BIT
	%%% WHEN YOU TALK MENTION THAT THEY DID THIS SURVEY AFTER PLAYING THE GAME IN CLASS
	\begin{frame}{Collecting Student Feedback}
		\begin{columns}
			\begin{column}{0.45\textwidth}
				\includegraphics[width=5cm, height=2cm]{nsse.png}
			\end{column}
			\begin{column}{0.5\textwidth}
				\begin{itemize}
					\item NSSE survey questions assess the extent to which students engage in educational practices in the context of higher levels of learning \newline
					\item 150 participants enrolled in BIO1A03 were asked 14 questions regarding opinions and attitudes on their current undergraduate education, as well as feedback on the use of `Cells at War' for learning in the biology classroom
				\end{itemize}
			\end{column}
		\end{columns}

	\end{frame}

	%%%%RESULTS BEGIN HERE
	\section{Results}

	\begin{frame}{How Much Time Students Spend Playing Video Games}
		\begin{center}
		\includegraphics[width=11cm, height=7cm]{time_spent_playing_videogames.jpg}
		\end{center}
	\end{frame}

	\begin{frame}{Types of Courses Taken at McMaster University}
		\begin{center}
		\includegraphics[width=11cm, height=6cm]{types_of_courses_taken.jpg}
		\end{center}
	\end{frame}

	\begin{frame}{During the Current School Year, How Often Have You Done the Following}
		\begin{center}
			\includegraphics[width=11cm, height=6cm]{how_often_haveyou_done_thefollowing.jpg}
		\end{center}
	\end{frame}

	\begin{frame}{How Could Game-based Learning Help With the Following}
		\begin{center}
			\includegraphics[width=11cm, height=6cm]{how_does_gbl_help_in_science.jpg}
		\end{center}
	\end{frame}

	\begin{frame}{How Much Does Coursework Emphasize the Following}
		\begin{center}
			\includegraphics[width=11cm, height=6cm]{howmuch_coursework_emphasized_thefollowing.jpg}
		\end{center}
	\end{frame}

	\begin{frame}{If BIO1A03 Added a GBL Component, How Would it Improve Motivation to do the Following}
		\begin{center}
			\includegraphics[width=11cm, height=6cm]{ifbio1a03_added_gblcomponent.jpg}
		\end{center}
	\end{frame}

	\begin{frame}{How Likely is it You Would Play Cells at War on Your Own Time To Consolidate Material Taught During Class}
		\begin{center}
			\includegraphics[width=11cm, height=6cm]{how_likely_you_would_play_cellsatwar.jpg}
		\end{center}
	\end{frame}

	\begin{frame}{How Prepared Would you Feel if Given a Quiz on Pompe Disease based on Cells at War, Compared to Studying off Traditional Lecture Slides}
		\begin{center}
			\includegraphics[width=11cm, height=6cm]{how_prepared_if_tested_on_pompe_based_off_cellsatwar.jpg}
		\end{center}
	\end{frame}

	\section{Conclusion}
	\begin{frame}{Conclusions}
		\begin{enumerate}	
			\item Game based learning approaches have the potential to be utilized during lecture as interactive educational tools for students. \newline
			\item Implementing Game-based learning practices could improve student motivation to apply and analyze what they learn, and use it to form new understandings of concepts. \newline
			\item In a cohort where many students did not play video games, learning to play did not distract from understanding the content. \newline
			\item Not only benefits for the students playing the game, but also for the team of students and supervisors from different schools, in different programs, working in collaboration with eachother.
		\end{enumerate}
	\end{frame}

	\section{Future Work}
	\begin{frame}{Future Work}
		\begin{itemize}
			\item Develop games based on different diseases to add to the Cells at War Suite of Games
			\item Expand the use of video games to other domains across STEAM (Physics, Music, Etc...)
			\item International collaboration between more Universities and Colleges	
		\end{itemize}
	\begin{columns}
		\begin{column}{0.4\textwidth}
			\movie[width=5cm,height=3cm,showcontrols=true]{\includegraphics[width=5cm, height=3cm]{tid_thumbnail.png}}{biodemonstration.mp4}
		\end{column}
		\begin{column}{0.3\textwidth}
			\includegraphics[width=3cm,height=3cm]{wgong.jpeg}
		\end{column}
	\end{columns}

	\end{frame}

	\section{Acknowledgements}
	\begin{frame}{Acknowledgements}
		Thank you...
		\begin{itemize}
			\item Dr. Rosa da Silva (McMaster)
			\item Jean Paul Amore (GBC)
			\item Bianca Flaim

		\end{itemize}
	\centering
	\includegraphics[width=5cm, height=3cm]{gbc.png}	
	\includegraphics[width=10cm, height=3cm]{cewil.jpg}
	\end{frame}

\end{document}

### 2023-04-02 thesis_template/paper0904.tex
Results section done, only discussion now

@@
 	\includegraphics[width=\textwidth]{figures_4f06/types_of_courses_taken.jpg}
-	\caption{Types of Courses Taken}
+	\caption{Student responses regarding the different types of courses taken at McMaster University in 2022}
 	\label{fig:2}
@@
 Participants responded to two questions regarding how much coursework has emphasized certain aspects of their learning \figref{fig:3}, and how often they have applied their learning in different contexts \figref{fig:4}.
-62
+The majority of students responded that their coursework emphasized applying facts to new problems, analyzing ideas in depth, and forming new ideas from various pieces of information "quite a bit".
+In the case of memorizing course material however, 41\% of participants responded this was emphasized "very much" in their courses.
 
@@
 	\includegraphics[width=\textwidth]{figures_4f06/howmuch_coursework_emphasized_thefollowing.jpg}
-	\caption{How Much Has Coursework Emphasized the following}
+	\caption{Student responses regarding how much their coursework has emphasized the following aspects}
 	\label{fig:3}
 \end{figure}
+
+\figref{fig:4} follows a similar format to the previous question.
+Students were asked how often they had combined ideas from different courses when completing assignments, connected their learning to societal issues, learned something that changed understanding of an issue and connected ideas from courses to prior experiences and knowledge.
+Most students responded doing these things "sometimes" or "often", with a smaller number of participants responding "very often".
+It is worthwhile to note that 43\% of students indicated that they "sometimes" connect what they have learned to societal or health related issues.
+As science students, it is important to connect what has been learned to societal and health related issues, as the concepts we study are interwoven throughout all facets of life.
+Implementing game based learning into science classrooms has the potential to change this narrative by presenting concepts to students in the form of real world examples, rather than static textbook diagrams.
 
@@
 	\includegraphics[width=\textwidth]{figures_4f06/how_often_haveyou_done_thefollowing.jpg}
-	\caption{How often have you done the following}
+	\caption{Student responses regarding how often they have done the following things during the school year}
 	\label{fig:4}
@@
 
+After the questions regarding current undergraduate experiences, students were asked to reflect on their feelings towards implementing game-based learning approaches into their classrooms.
+
+Following the same question structure as \figref{fig:4}, students were asked how implementing game-based learning into any one of their science courses could help in a wide variety of areas.
+Interestingly, across all four areas, we see an increase in the number of students who responded "very much", compared to the number of students who responded "very much" in \figref{fig:4}.
+Through implementing game based learning methods in science classrooms, more students would feel better suited to combine ideas from different courses when completing assignments, connect their learning to societal and health related issues and connect ideas from courses to prior experiences and knowledge.
+Following the same question structure as \figref{fig:3}, \figref{fig:6} shows how motivation to do certain things can be improved in BIO1A03 if a game-based learning component was added.
+In general, students indicated that a game-based learning component in BIO1A03 would improve motivation to connect ideas to societal issues, learn something that changes understanding and connect ideas from courses to prior experiences and knowledge "very much".
+In comparison to \figref{fig:3}, there is a clear increase in the number of responses which indicate "very much".
+The only exception however is in terms of motivation to memorize course material.
+There was a decrease in the number of students who reported that if BIO1A03 added a game-based learning component, it would motivate them to memorize course material.
+This is exactly the aim of game-based learning approaches, to emphasize exploration and understanding, not memory.
+
 %%%%%If game-based learning approaches were incorporated into one of your science courses, how do you think that this mode of learning can help with the following
@@
 	\includegraphics[width=\textwidth]{figures_4f06/how_does_gbl_help_in_science.jpg}
-	\caption{If game-based learning approaches were incorporated in your science courses}
+	\caption{Student responses regarding how game-based learning approaches could help with the following things if implemented in science courses}
 	\label{fig:5}
@@
 	\includegraphics[width=\textwidth]{figures_4f06/ifbio1a03_added_gblcomponent.jpg}
-	\caption{If BIO1A03 added a game-based component}
+	\caption{Student responses regarding how motivation to do the following things would improve if BIO1A03 added a game-based learning component}
 	\label{fig:6}
 \end{figure}
+
+The final two questions provided insight into students' feelings towards not only game-based learning, but specifically towards Cells at War.
+Students were asked how likely it is that they would play Cells at War outside of class on their own time to further consolidate material taught during class \figref{fig:7}.
+The results indicated that 43\% of students were somewhat likely to do this, while about 17\% of students reported being very likely to this.
+21\% of participants indicated they would not be very likely to do this.
+These results highlight the potential game-based learning has to be utilized not only as an instructional tool in class, but also as a form of practice and content mastery for students.
 
@@
 	\includegraphics[width=\textwidth]{figures_4f06/how_likely_you_would_play_cellsatwar.jpg}
-	\caption{How likely is it you would play Cells at War outside of class}
+	\caption{Student responses regarding how likely they would be to play Cells at War on their own time, outside of class to further consolidate material taught during class}
 	\label{fig:7}
 \end{figure}
+
+To assess if students understood the content shown to them in game, they were asked how prepared they would feel if given a quiz on Pompe Disease based off playing Cells at War, compared to studying off traditional lecture slides \figref{fig:8}.
+52\% of students indicated they would be somewhat prepared, and 17\% of students indicated they would feel very prepared.
+This strengthens the argument that again, game-based learning is not only an in-class intervention.
+Video games promote practice, critical thinking, and content mastery, all of which are important aspects when preparing for assessment.
+Students are willing to try alternative studying methods, and giving them the ability to study in a way that is engaging and interacting could have positive benefits on performance.
 
@@
 	\includegraphics[width=\textwidth]{figures_4f06/how_prepared_if_tested_on_pompe_based_off_cellsatwar.jpg}
-	\caption{How prepared would you feel if given a quiz on Pompe Disease}
+	\caption{Student responses regarding how prepared they would feel if given a quiz on Pompe Disease based off Cells at War compared to studying off traditional lecture slides}
 	\label{fig:8}

### 2023-04-04 thesis_template/paper0904.tex
started writing discussion

@@
-\documentclass{article}
+\documentclass[10pt]{article}
 %\documentclass[twocolumn]{article}
@@
 Students pitched these proposals to supervising professors who then selected the ideas and game elements that best suited the goals of the project.
+The disease that was selected was Pompe disease, a rare genetic disease that causes skeletal and cardiac muscle weakness due to the accumulation of glycogen within lysosomes.
+The game envisioned was a shooter type game, where students would be tasked with shooting enzymes to break down glycogen into glucose in order to remove glycogen from the lysosomes.
+Each level highlighted a different stage of the disease; a healthy individual, early onset pompe disease individual, and full fledged pompe disease individual.
+Through each level, students would be able to see the effects of having less enzymes to break down glycogen.
 
@@
 MORE HERE ABOUT NSSE...
+\newpage
 
@@
 
+\newpage
+
+%We need to highlight in this section the use of Cells at War as an instructional tool as well as the benefits that students get out of this work integrated learning experience
 \section{Discussion}
 
-We need to highlight in this section the use of Cells at War as an instructional tool as well as the benefits that students get out of this work integrated learning experience
+This project has developed an educational video game focused on a single gene disease, which presents students with important cellular and molecular biology concepts in an engaging and interactive manner.
+Through this study, we have found a way to supplement lecture material with a video game, and examined the impact of game-based learning on student motivation and engagement.
+This study also shows the immense benefits both students and faculty gain when engaging in a collaborative and interdisciplinary work-integrated learning experience.
+
+\subsection{Student benefits of game-based learning}
+
+\subsection{Benefits of the game-design process for educators and students}
 

### 2023-04-05 cellsatwar_bus_presentation/beap_pres_dec.tex
Updated pres

@@
 \author{Alexander Turco}
-\date{December 5, 2022}
+\date{April 5, 2023}
 \logo{\includegraphics[height=1cm, width=1cm]{logo.png}}
@@
 			\item In a cohort where many students did not play video games, learning to play did not distract from understanding the content. \newline
-			\item Not only benefits for the students playing the game, but also for the team of students and supervisors from different schools, in different programs, working in collaboration with eachother.
+			\item Not only benefits for the students playing the game, but also for the team of students and supervisors from different schools, in different programs, working in collaboration with each other.
 		\end{enumerate}

### 2023-04-05 cellsatwar_bus_presentation/beap_pres_dec.tex
BUS PRES DONE

@@
 \author{Alexander Turco}
-\date{April 5, 2023}
+\date{April 13, 2023}
 \logo{\includegraphics[height=1cm, width=1cm]{logo.png}}
@@
 	% Outline frame - SLIDE 2
-	\begin{frame}{Overview}
+	%\begin{frame}{Overview}
 
-		\begin{center}
-		\begin{minipage}{6cm}
+		%\begin{center}
+		%\begin{minipage}{6cm}
 
-		  		\begin{block}{} \hyperlink{link1}{Background Information} \end{block}
-		  		\begin{block}{} The Game Design Process \end{block}
-		  		\begin{block}{} Collection of Student Feedback \end{block}
-		  		\begin{block}{} Results \end{block}
-		  		\begin{block}{} Conclusions \end{block}
-		  		\begin{block}{} Future Work \end{block}
+		 % 		\begin{block}{} \hyperlink{link1}{Background Information} \end{block}
+		  %		\begin{block}{} The Game Design Process \end{block}
+		  %		\begin{block}{} Collection of Student Feedback \end{block}
+		  %		\begin{block}{} Results \end{block}
+		  		%\begin{block}{} Conclusions \end{block}
+		  %		\begin{block}{} Future Work \end{block}
 
-		\end{minipage}
-		\end{center}
+		%\end{minipage}
+		%\end{center}
 
-	\end{frame}
+	%\end{frame}
 
@@
 		\includegraphics[width=11cm, height=7cm]{time_spent_playing_videogames.jpg}
+		\begin{itemize}
+			\item 44\% of students reported not playing video games at all
+		\end{itemize}
 		\end{center}
@@
 			\includegraphics[width=11cm, height=6cm]{how_does_gbl_help_in_science.jpg}
+			\begin{itemize}
+				\item 22\% increase in the number of respondants who felt that game-based learning would help connect their learning to societal/health related issues
+			\end{itemize}
 		\end{center}
@@
 			\includegraphics[width=11cm, height=6cm]{ifbio1a03_added_gblcomponent.jpg}
+			\begin{itemize}
+				\item Overall increase in the number of students responding "very much" in all cases except memorizing course material
+			\end{itemize}
 		\end{center}
@@
 	\begin{frame}{Acknowledgements}
+		\begin{columns}
+			\begin{column}{0.45\textwidth}
 				Thank you...
 				\begin{itemize}
-			\item Dr. Rosa da Silva (McMaster)
-			\item Jean Paul Amore (GBC)
+					\item Dr. Rosa da Silva
+					\item Jean Paul Amore
 					\item Bianca Flaim
+					\item Oluwalonimi Ayekun
+					\item Jennifer Luk
+					\item The Entire George Brown College team
+					\item CEWIL Canada	
+				\end{itemize}
+			\end{column}
+			\begin{column}{0.5\textwidth}
+				\centering
+				\includegraphics[width=2cm, height=1cm]{gbc.png}	
+				\includegraphics[width=5cm, height=2cm]{cewil.jpg}
+			\end{column}
+		\end{columns}
 
-		\end{itemize}
-	\centering
-	\includegraphics[width=5cm, height=3cm]{gbc.png}	
-	\includegraphics[width=10cm, height=3cm]{cewil.jpg}
 	\end{frame}

### 2023-04-10 thesis_template/paper0904.tex
Updates to discussion

@@
 This project has developed an educational video game focused on a single gene disease, which presents students with important cellular and molecular biology concepts in an engaging and interactive manner.
-Through this study, we have found a way to supplement lecture material with a video game, and examined the impact of game-based learning on student motivation and engagement.
-This study also shows the immense benefits both students and faculty gain when engaging in a collaborative and interdisciplinary work-integrated learning experience.
+Through this study, we have found a way to enhance the delivery of lecture material with an interactive video game, and examined the impact of game-based learning on student motivation and engagement.
+This study also highlights the immense benefits both students and faculty gain when engaging in a collaborative and interdisciplinary work-integrated learning experience.
 
@@
 
+Video games have been around for over 30 years, however, the rate at which technology is evolving has allowed video games to change from short, frenetic, and repetitive forms of engagement, to long, descriptive narratives and story lines \citep{annetta2008video}.
+Although the results of this study lack quantitative data to support the use of video games in University classrooms, the qualitative feedback provided by students emphasizes the massive potential game-based learning interventions could have at the post-secondary level.
+\citet{o2009garage} report that static images used in lecture slides are not seemingly effective in conveying dynamic aspects of biological processes.
+The results from our study provide evidence for this as 69\% of students reported being either somewhat or very prepared if they were to be tested on material using the video game to study, as opposed to studying off of traditional lecture slides.
+This highlights the need to shift from old fashioned, didactic educational practices, to newer methods that focus on keeping students engaged with the material being presented.
+
+It has also been found that creating a more learner-centered environment in a large introductory biology course by replacing certain material with active learning exercises, can potentially increase engagement, promote critical thinking, and improve student attitudes \citep{moravec2010learn, o2009garage}.
+Students were asked to report how game based learning interventions could help with following four things; combining ideas from different courses when completing assignments, connecting learning to societal/health related issues, connecting ideas from courses to prior experiences and knowledge, and learning something that changes understanding of a concept.
+The survey results show an overall increase in the number of students who responded "very much" in all four of these categories, compared to when they were asked how much they were currently doing these things without game-based learning.
+Video-games have the potential to be powerful educational tools that allow students to connect what they learn through engaging game dynamics, to other aspects of their lives.
+
+It is also interesting to note that the survey results indicated that 41\% of students felt that their coursework emphasized memorizing course material "very much".
+When asked how a game-based learning intervention could improve motivation to memorize course material, the results show a 14\% decrease in the number of students who responded "very much".
+This highlights the ability video games have to promote understanding through play, rather than memorization of static textbook and lecture content.
+
 \subsection{Benefits of the game-design process for educators and students}
+
+Aside from the benefits that game-based learning has for students utilizing the tool, the educators and students involved in the game design also reap massive benefits.
+There has been a rising interest in the realm of higher education about "students as partners" and co-created learning and teaching \citep{bibid}
 

### 2023-04-12 thesis_template/paper0904.tex
DIscussion for paper finished

@@
 
-Aside from the benefits that game-based learning has for students utilizing the tool, the educators and students involved in the game design also reap massive benefits.
-There has been a rising interest in the realm of higher education about "students as partners" and co-created learning and teaching \citep{bibid}
+The Cells at War project was rooted in an interdisciplinary approach in which biology students, artists, programmers and game designers came together to share their expertise in their related fields.
+This approach benefited students in many ways, one of which was having students move from a more teacher centred approach to learning, to an instructional approach in which the students and educators co-create and co-learn together across a variety of disciplines.
+Benjamin Franklin’s old adage of “Tell me and I forget, teach me and I may remember, involve me and I learn” seems to have some merit as what the project served was to foster a learning environment in which students become empowered, engaged and self-motivated when working and learning alongside their teachers.
+Universities are typically structured in discipline-centred silos, which encourage discipline specific research.
+Many benefits arise however when individuals from a variety of disciplines come together \citep{snow201026,martin2020value}.
+Interdisciplinary teams are effective when they address complex, real-life problems and incorporate different perspectives \citep{beers2006common, choi2007multidisciplinarity, fischer2005knowledge, massey2006crossing, martin2020value}.
+Further, interdisciplinary projects encourage face-to-face interactions, develop a shared understanding of the problems they are attempting to address, and achieve effective team dynamics by building trusting, open environments that are not based on hierarchy \citep{stokols2005vivo, fischer2005knowledge, bruusgaard2010we, martin2020value}.
+Cells at War is a game based learning project in which students engage in experiential learning.
+Students working in interdisciplinary settings are reported to benefit from experiential learning as this leads to critical self-reflection, the development of transferable skills, and contributes to new ideas \citep{gemmel2009comprehensive, martin2020value}.
+Students at McMaster University and George Brown College were involved in a symbiotic process in which educators and students with different perspectives, abilities, and ideas were able to combine and apply their strengths and diverse knowledge and skills to build an educational video game that served as an engaging learning tool.
 
-\section{Future Directions}
+Aside from the benefits of engaging in an interdisciplinary approach, Cells at War is a highly engaging instructional tool developed by students, for students.
+The success of the development of the video game rests in educators and students working together in the game design process.
+This collaborative approach reaps mutual benefits and rewards.
+There has been a rising interest in the realm of higher education about "students as partners" and co-created learning and teaching \citep{felten2014engaging, dunne2016design, mercer2017systematic, bovill2020co}.
+As pedagogy in higher education has evolved, co-creation of learning has created more inclusive classrooms where students are able to build positive relationships with teachers, and students with students \citep{bovill2020co}.
+As the McMaster and George Brown team engaged in the development process of the video game, what emerged was a strong student voice in which students had a say in what they were learning and how to approach it.
+The dialogue that emerged and the exchange of ideas not only served to create an innovative instructional tool but, reinforced and consolidated the many skills and concepts that students from each discipline were required to learn.
+Greater student-teacher interaction leads to students’ academic success, good academic performance, higher educational aspirations, personal and intellectual development, student satisfaction, and enhanced motivation \citep{bovill2020co}.
+
+The benefits to students engaging in the learning process are many.
+Of equal importance is the benefit for educators engaged in this process.
+Using novel approaches to improve student learning is an ongoing consideration for all educators.
+The process in which Cells at War was developed, provided teachers with an alternate way to engage and motivate students through best practices.
+In addition, teachers were able to provide feedback to students, which also supported them in informing their next point of instruction.
+
+%\section{Future Directions}
 

### 2023-06-07 cellsatwar_bus_presentation/beap_pres_dec.tex
BUS PRESENTATION READY TO GO
updates for western conference

@@
 % Bibliography stuff
-\usepackage[natbib=true, sorting=nyt, style=authoryear-comp]{biblatex}
-\addbibresource{ABC1.bib}
-
+\usepackage[backend=biber, giveninits=true, doi=false, isbn=false, natbib=true, url=true, eprint=false, style=numeric, sorting=nyt, sortcites=ynt, maxcitenames=2, maxbibnames=10, minbibnames = 10, uniquename=false, uniquelist=false]{biblatex} % can change the maxbibnames to cut long author lists to specified length followed by et al., currently set to 99.
+\addbibresource{ABC1.bib} % The filename of the bibliography
 %Extra packages
@@
 
-			\footnotetext[1]{\tiny\cite{brown1989situated, jan2016game}}
+		\tiny\footcite{brown1989situated, jan2016game}
 		\end{itemize}
@@
 
-	\footnotetext[2]{\tiny\cite{coller2009effectiveness, tomavska2022let}}
+	{\tiny\footcite{coller2009effectiveness, tomavska2022let}}
 
@@
 
-	\footnotetext[3]{\tiny\cite{lister2015gamification,mayer2010adding,molin2017role,fishman2014empowering}}
+	{\tiny\footcite{lister2015gamification,mayer2010adding,molin2017role,fishman2014empowering}}
 
@@
 				\centering
+				\includegraphics[width=4cm, height=5cm]{credits.png}
 				\includegraphics[width=2cm, height=1cm]{gbc.png}	
-				\includegraphics[width=5cm, height=2cm]{cewil.jpg}
+				\includegraphics[width=4cm, height=1cm]{cewil.jpg}
 			\end{column}
 		\end{columns}
+	\end{frame}
 
+	\section{References}
+	\begin{frame}[allowframebreaks]
+		\frametitle{References}
+		\nocite{*}
+		\printbibliography[heading=none, sorting=nyt]
+		\newpage
 	\end{frame}

### 2023-09-26 survey_stuff_2022_cellsatwar_foranalysis/survey_analysis.R
updates

@@ -5,6 +5,7 @@ morning_df_csv <- read.csv("Depersonalized_survey_data_morning(2).csv", header =
 combined_df_csv <- read.csv("combined_data_morning_afternoon_raw_CSV.csv", header = TRUE)
 
 ###CREATING SOME GRAPHS
+library(dplyr)
 library(ggplot2)
 library(ggthemes)
 library(ggpubr) #For ggarrange function
@@ -12,15 +13,17 @@ library(ggpubr) #For ggarrange function
 ### FUNCTION FOR PUMPING OUT BARPLOTS
 plot_survey_fun <- function(df, bar_lables, xtitle, ytitle, chart_title) {
   p <- ggplot(df, aes(x = df[,1], y = df[,2])) + 
-    geom_col(show.legend = FALSE, width = 0.7, fill = "#000075") +
-    geom_text(aes(label = df[,2]), vjust = -0.5, colour = "black", nudge_x = -0.14 ) +
+    geom_bar(show.legend = FALSE, width = 0.7, fill = "#000075", stat = "identity") +
+    geom_text(aes(label = df[,2]), vjust = -0.5, colour = "black", nudge_x = -0.14) +
     geom_text(aes(label = paste0("(",round(df[,2]/sum(df[,2])*100),")%")),vjust = -0.5, nudge_x = 0.14 ) +
+    geom_text(x = -Inf, y = Inf, label = paste0("n = ", sum(df[,2])), 
+              hjust = -1, vjust = 3, size = 4, color = "black") +
     scale_x_discrete(labels = bar_lables) +
     xlab(xtitle) +
     ylab(ytitle) +
     ggtitle(chart_title)
   
-  p + theme_clean(base_size = 8)
+  p + theme_clean(base_size = 11.5)
 }
 
 #Q1
@@ -29,38 +32,30 @@ plot_survey_fun <- function(df, bar_lables, xtitle, ytitle, chart_title) {
 ########################################################################
 
 time_spent_vg <- as.data.frame(table(combined_df_csv$On.average..how.much.time.do.you.spend.playing.video.games.each.week....Select.one.option.))
-time_spent_vg_labels <- c("NA", "I don't play video games", "Between 10-20 hours/week", "3-10 hours/week", "Under 3 hours a week", "More than 20 hours/week")
[…]
